---
title: "Pathways Persistence of MDROs Analysis"
author: "Skylar Zeng"
date: "Last compiled on `r format(Sys.Date(), '%B %d, %Y')`"
format:
  docx:
    code-overflow: wrap
embed-resources: true
---

# Install and Load Libraries
```{r}
#| label: install_packages
#| eval: false
install.packages("readr")
install.packages("readxl")
install.packages("dplyr")
install.packages("tidyr")
install.packages("gtsummary")
install.packages("gt")
install.packages("sjlabelled")
install.packages("writexl")
install.packages("parameters")
install.packages("bayestestR")
install.packages("purrr")
install.packages("ggplot2")
install.packages("ggpubr")
install.packages("survival")
install.packages("ranger")
install.packages("ggfortify")
install.packages("survminer")
install.packages("lmtest")
install.packages("labelled")
```

```{r}
#| label: load_packages
#| message: false
library(readr)
library(readxl)
library(dplyr)
library(tidyr)
library(gtsummary)
library(gt)
library(sjlabelled)
library(writexl)
library(parameters)
library(bayestestR)
library(purrr)
library(ggplot2)
library(ggpubr)
library(survival)
library(ranger)
library(ggfortify)
library(survminer)
library(lmtest)
library(labelled)
```

# Import Datasets
```{r}
#| label: import_dataset
#| message: false
Pathways_complete <- read_csv("Pathways_both phases final May 2021 wound update.csv")
Pathways_dictionary <- read_excel("Pathways Database Dictionary_Derived Vars in second sheet.xlsx")
# New Acquisition
rgnb_acquisition_persistence <- read_excel("rgnb_acquisition_persistence_patientlevel.xlsx")
mrsa_acquisition_persistence <- read_excel("mrsa_acquisition_persistence_patientlevel.xlsx")
vre_acquisition_persistence <- read_excel("vre_acquisition_persistence_patientlevel.xlsx")
# Colonization
rgnb_colonization_persistence <- read_excel("gnbr_colonization_persistence_patientlevel.xlsx")
mrsa_colonization_persistence <- read_excel("mrsa_colonization_persistence_patientlevel.xlsx")
vre_colonization_persistence <- read_excel("vre_colonization_persistence_patientlevel.xlsx")
```
## Subsets with Necessary Variables
```{r}
test_subset <- Pathways_complete[, c("facilID", 
                                    "studyID",
                                    "visit", 
                                    "novisits", 
                                    "totvisits", 
                                    "phase", 
                                    "maxvisit",
                                    "age_bl", 
                                    "sex_bl",
                                    "height_bl",
                                    "weight_bl",
                                    "race_bl",
                                    "ethnic_bl",
                                    "dtvisit", 
                                    "intervention",
                                    "antibiotic_bl", 
                                    "feedtube", 
                                    "urincath", 
                                    "deviceuse", 
                                    "openwound_final",
                                    "cogsum_bl",
                                    "homrsa_bl",
                                    "hovre_bl",
                                    "horgnb_bl",
                                    "psmscore",
                                    "anyanypos_p",
                                    "anymrsa_p",
                                    "anyvre_p",
                                    "anygnbr_p",
                                    "acqmrsa", 
                                    "acqvre", 
                                    "acqgnbr", 
                                    "hospstay", 
                                    "charlson", 
                                    "ccsmi_bl", 
                                    "ccschf_bl", 
                                    "ccspvod_bl", 
                                    "ccscva_bl", 
                                    "ccsdemen_bl", 
                                    "ccscopd_bl", 
                                    "ccscontis_bl", 
                                    "ccspud_bl", 
                                    "ccsmld_bl", 
                                    "ccsdm_bl", 
                                    "ccshemi_bl", 
                                    "ccsmsrd_bl", 
                                    "ccsdmeod_bl", 
                                    "ccstumor_bl", 
                                    "ccsleuk_bl", 
                                    "ccslymph_bl", 
                                    "ccsmsld_bl", 
                                    "ccsmst_bl",
                                    "ccsaids_bl",
                                    "VREsiteE1",
                                    "VREsiteE10",
                                    "VREsiteE11",
                                    "VREsiteE12",
                                    "VREsiteE2",
                                    "VREsiteE3", 
                                    "VREsiteE4",
                                    "VREsiteE5",
                                    "VREsiteE6",
                                    "VREsiteE7",
                                    "VREsiteE8",
                                    "VREsiteE9",
                                    "VREsiteF",
                                    "VREsiteG",
                                    "VREsiteH",
                                    "VREsiteN",
                                    "VREsiteO",
                                    "VREsiteR",
                                    "VREsiteS",
                                    "vre_p",
                                    "MRSAsiteE1",
                                    "MRSAsiteE10",
                                    "MRSAsiteE11",
                                    "MRSAsiteE12",
                                    "MRSAsiteE2",
                                    "MRSAsiteE3", 
                                    "MRSAsiteE4",
                                    "MRSAsiteE5",
                                    "MRSAsiteE6",
                                    "MRSAsiteE7",
                                    "MRSAsiteE8",
                                    "MRSAsiteE9",
                                    "MRSAsiteF",
                                    "MRSAsiteG",
                                    "MRSAsiteH",
                                    "MRSAsiteN",
                                    "MRSAsiteO",
                                    "MRSAsiteR",
                                    "MRSAsiteS",
                                    "mrsa_p",
                                    "RGNBsiteE1",
                                    "RGNBsiteE10",
                                    "RGNBsiteE11",
                                    "RGNBsiteE12",
                                    "RGNBsiteE2",
                                    "RGNBsiteE3",
                                    "RGNBsiteE4",
                                    "RGNBsiteE5",
                                    "RGNBsiteE6",
                                    "RGNBsiteE7",
                                    "RGNBsiteE8",
                                    "RGNBsiteE9",
                                    "RGNBsiteF",
                                    "RGNBsiteG",
                                    "RGNBsiteH",
                                    "RGNBsiteN",
                                    "RGNBsiteO",
                                    "RGNBsiteR",
                                    "RGNBsiteS",
                                    "gnbr_p",
                                    "gnbr_env",
                                    "mrsa_env",
                                    "vre_env"
                                    )]

# Replace "99" with NA in all columns
test_subset[test_subset == "99"] <- NA
test_subset[test_subset == "999"] <- NA

# Create BMI variable
test_subset <- test_subset %>%
  mutate(BMI = (weight_bl * 703) / (height_bl^2))

# Create visit_intervened variable (Intervention arm are all pnts in facility 1,2,6; interventions occurred in Phase 2)
test_subset <- test_subset %>%
  mutate(visit_intervened = case_when(
    phase == 1 ~ 0,
    phase == 2 & intervention == 0 ~ 0,
    phase == 2 & intervention == 1 ~ 1,
  ))

# Create new variable - gender
test_subset <- test_subset %>% 
  mutate(gender = case_when(sex_bl == 1 ~ "Male",
                            sex_bl == 0 ~ "Female"))
# Create new variable - race
test_subset <- test_subset %>% 
  mutate(race_bl = case_when(race_bl == 1 ~ 0,
                             race_bl == 2 ~ 1))

test_subset <- test_subset %>% 
  mutate(race = case_when(race_bl == 0 ~ "White",
                          race_bl == 1 ~ "Black or African American"))

# Create new categorical variable for charlson score
test_subset$charlson_cat <- ifelse(test_subset$charlson >= 5, ">=5", as.character(test_subset$charlson))

# Convert to factor with proper order
test_subset$charlson_cat <- factor(
  test_subset$charlson_cat, 
  levels = c("0", "1", "2", "3", "4", ">=5"),
  ordered = TRUE
)

# Create new categorical variable for Body Sites colonized
test_subset$gnbr_p_cat <- ifelse(test_subset$gnbr_p >= 3, ">=3", as.character(test_subset$gnbr_p))
test_subset$mrsa_p_cat <- ifelse(test_subset$mrsa_p >= 3, ">=3", as.character(test_subset$mrsa_p))
test_subset$vre_p_cat <- ifelse(test_subset$vre_p >= 3, ">=3", as.character(test_subset$vre_p))

# Convert to factor with proper order
test_subset$gnbr_p_cat <- factor(
  test_subset$gnbr_p_cat, 
  levels = c("0", "1", "2", ">=3"),
  ordered = TRUE
)
test_subset$mrsa_p_cat <- factor(
  test_subset$mrsa_p_cat, 
  levels = c("0", "1", "2", ">=3"),
  ordered = TRUE
)
test_subset$vre_p_cat <- factor(
  test_subset$vre_p_cat, 
  levels = c("0", "1", "2", ">=3"),
  ordered = TRUE
)

# Create new categorical variable for Body Sites colonized
test_subset$gnbr_env_cat <- ifelse(test_subset$gnbr_env >= 4, ">=4", as.character(test_subset$gnbr_env))
test_subset$mrsa_env_cat <- ifelse(test_subset$mrsa_env >= 4, ">=4", as.character(test_subset$mrsa_env))
test_subset$vre_env_cat <- ifelse(test_subset$vre_env >= 4, ">=4", as.character(test_subset$vre_env))

# Convert to factor with proper order
test_subset$gnbr_env_cat <- factor(
  test_subset$gnbr_env_cat, 
  levels = c("0", "1", "2", "3", ">=4"),
  ordered = TRUE
)
test_subset$mrsa_env_cat <- factor(
  test_subset$mrsa_env_cat, 
  levels = c("0", "1", "2", "3", ">=4"),
  ordered = TRUE
)
test_subset$vre_env_cat <- factor(
  test_subset$vre_env_cat, 
  levels = c("0", "1", "2", "3", ">=4"),
  ordered = TRUE
)

# Create a new variable for presence of comorbidities
test_subset <- test_subset %>%
  mutate(comorbidity_category = case_when(
    ccsmi_bl == 1 ~ "Myocardial Infarction",
    ccschf_bl == 1 ~ "Congestive Heart Failure",
    ccspvod_bl == 1 ~ "Peripheral Vascular Dx",
    ccscva_bl == 1 ~ "Cerebral Vascular Accident",
    ccsdemen_bl == 1 ~ "Dementia",
    ccscopd_bl == 1 ~ "COPD",
    ccscontis_bl == 1 ~ "Connective Tissue Dx",
    ccspud_bl == 1 ~ "Peptic Ulcer Dx",
    ccsmld_bl == 1 ~ "Mild Liver Dx",
    ccsdm_bl == 1 ~ "Diabetes",
    ccshemi_bl == 1 ~ "Hemiplegia",
    ccsmsrd_bl == 1 ~ "Mod-Severe Renal Dx",
    ccsdmeod_bl == 1 ~ "Diabetes End Organ Dx",
    ccstumor_bl == 1 ~ "Any Tumor",
    ccsleuk_bl == 1 ~ "Leukemia",
    ccslymph_bl == 1 ~ "Lymphoma",
    ccsmst_bl == 1 ~ "Metastatic Solid Tumor",
    ccsaids_bl == 1 ~ "AIDS",
    TRUE ~ "None"
  ))

# Set labels for the specified variables
test_subset <- test_subset %>%
  set_variable_labels(
    age_bl = "Age",
    gender = "Sex",
    race = "Race",
    antibiotic_bl = "Antibiotic Use",
    charlson = "Charlson comorbidity index score",
    charlson_cat = "Charlson comorbidity index score",
    feedtube = "Feeding Tube Use",
    urincath = "Urinary Catheter Use",
    deviceuse = "Device use",
    cogsum_bl = "Cognitive Sum Score",
    homrsa_bl = "History of MRSA",
    hovre_bl = "History of VRE",
    horgnb_bl = "History of RGNB",
    psmscore = "Functional disability (PSMS Total Score)",
    hospstay = "Length of Hospitalization",
    ccsmi_bl = "Myocardial Infarction",
    ccschf_bl = "Congestive Heart Failure",
    ccspvod_bl = "Peripheral Vascular Dx",
    ccscva_bl = "Cerebral Vascular Accident",
    ccsdemen_bl = "Dementia",
    ccscopd_bl = "COPD",
    ccscontis_bl = "Connective Tissue Dx",
    ccspud_bl = "Peptic Ulcer Dx",
    ccsmld_bl = "Mild Liver Dx",
    ccsdm_bl = "Diabetes",
    ccshemi_bl = "Hemiplegia",
    ccsmsrd_bl = "Mod-Severe Renal Dx",
    ccsdmeod_bl = "Diabetes End Organ Dx",
    ccstumor_bl = "Any Tumor",
    ccsleuk_bl = "Leukemia",
    ccslymph_bl = "Lymphoma",
    ccsmsld_bl = "Mod-Severe Liver Dx",
    ccsmst_bl = "Metastatic Solid Tumor",
    ccsaids_bl = "AIDS",
    openwound_final = "Wound",
    visit_intervened = "Intervention",
    comorbidity_category = "Presence of Comorbidities",
    gnbr_env_cat = "Number of Room Surfaces Colonized",
    mrsa_env_cat = "Number of Room Surfaces Colonized",
    vre_env_cat = "Number of Room Surfaces Colonized"
  )

test_subset_rgnb <- test_subset[, c("facilID", 
                                    "studyID",
                                    "visit", 
                                    "novisits", 
                                    "totvisits", 
                                    "phase", 
                                    "maxvisit",
                                    "age_bl", 
                                    "sex_bl",
                                    "height_bl",
                                    "weight_bl",
                                    "race_bl",
                                    "ethnic_bl",
                                    "dtvisit", 
                                    "intervention",
                                    "antibiotic_bl", 
                                    "feedtube", 
                                    "urincath", 
                                    "deviceuse", 
                                    "openwound_final",
                                    "cogsum_bl",
                                    "homrsa_bl",
                                    "hovre_bl",
                                    "horgnb_bl",
                                    "psmscore",
                                    "anyanypos_p",
                                    "anymrsa_p",
                                    "anyvre_p",
                                    "anygnbr_p",
                                    "acqmrsa", 
                                    "acqvre", 
                                    "acqgnbr", 
                                    "hospstay", 
                                    "charlson", 
                                    "ccsmi_bl", 
                                    "ccschf_bl", 
                                    "ccspvod_bl", 
                                    "ccscva_bl", 
                                    "ccsdemen_bl", 
                                    "ccscopd_bl", 
                                    "ccscontis_bl", 
                                    "ccspud_bl", 
                                    "ccsmld_bl", 
                                    "ccsdm_bl", 
                                    "ccshemi_bl", 
                                    "ccsmsrd_bl", 
                                    "ccsdmeod_bl", 
                                    "ccstumor_bl", 
                                    "ccsleuk_bl", 
                                    "ccslymph_bl", 
                                    "ccsmsld_bl",
                                    "ccsmst_bl", 
                                    "ccsaids_bl",
                                "comorbidity_category",
                                    "RGNBsiteE1",
                                    "RGNBsiteE10",
                                    "RGNBsiteE11",
                                    "RGNBsiteE12",
                                    "RGNBsiteE2",
                                    "RGNBsiteE3",
                                    "RGNBsiteE4",
                                    "RGNBsiteE5",
                                    "RGNBsiteE6",
                                    "RGNBsiteE7",
                                    "RGNBsiteE8",
                                    "RGNBsiteE9",
                                    "RGNBsiteF",
                                    "RGNBsiteG",
                                    "RGNBsiteH",
                                    "RGNBsiteN",
                                    "RGNBsiteO",
                                    "RGNBsiteR",
                                    "RGNBsiteS",
                                    "gnbr_p",
                                "gnbr_p_cat",
                                    "BMI",
                                    "visit_intervened",
                                    "gender",
                                    "race",
                                    "charlson_cat",
                                "gnbr_env",
                                    "gnbr_env_cat"
                                    )]

test_subset_mrsa <- test_subset[, c("facilID", 
                                    "studyID",
                                    "visit", 
                                    "novisits", 
                                    "totvisits", 
                                    "phase", 
                                    "maxvisit",
                                    "age_bl", 
                                    "sex_bl",
                                    "height_bl",
                                    "weight_bl",
                                    "race_bl",
                                    "ethnic_bl",
                                    "dtvisit", 
                                    "intervention",
                                    "antibiotic_bl", 
                                    "feedtube", 
                                    "urincath", 
                                    "deviceuse", 
                                    "openwound_final",
                                    "cogsum_bl",
                                    "homrsa_bl",
                                    "hovre_bl",
                                    "horgnb_bl",
                                    "psmscore",
                                    "anyanypos_p",
                                    "anymrsa_p",
                                    "anyvre_p",
                                    "anygnbr_p",
                                    "acqmrsa", 
                                    "acqvre", 
                                    "acqgnbr", 
                                    "hospstay", 
                                    "charlson", 
                                    "ccsmi_bl", 
                                    "ccschf_bl", 
                                    "ccspvod_bl", 
                                    "ccscva_bl", 
                                    "ccsdemen_bl", 
                                    "ccscopd_bl", 
                                    "ccscontis_bl", 
                                    "ccspud_bl", 
                                    "ccsmld_bl", 
                                    "ccsdm_bl", 
                                    "ccshemi_bl", 
                                    "ccsmsrd_bl", 
                                    "ccsdmeod_bl", 
                                    "ccstumor_bl", 
                                    "ccsleuk_bl", 
                                    "ccslymph_bl", 
                                    "ccsmsld_bl",
                                    "ccsmst_bl", 
                                    "ccsaids_bl",
                                "comorbidity_category",
                                    "MRSAsiteE1",
                                    "MRSAsiteE10",
                                    "MRSAsiteE11",
                                    "MRSAsiteE12",
                                    "MRSAsiteE2",
                                    "MRSAsiteE3", 
                                    "MRSAsiteE4",
                                    "MRSAsiteE5",
                                    "MRSAsiteE6",
                                    "MRSAsiteE7",
                                    "MRSAsiteE8",
                                    "MRSAsiteE9",
                                    "MRSAsiteF",
                                    "MRSAsiteG",
                                    "MRSAsiteH",
                                    "MRSAsiteN",
                                    "MRSAsiteO",
                                    "MRSAsiteR",
                                    "MRSAsiteS",
                                    "mrsa_p",
                                    "BMI",
                                    "visit_intervened",
                                    "gender",
                                    "race",
                                    "charlson_cat",
                                "mrsa_env",
                                "mrsa_p_cat",
                                "mrsa_env_cat"
                                    )]

test_subset_vre <- test_subset[, c("facilID", 
                                    "studyID",
                                    "visit", 
                                    "novisits", 
                                    "totvisits", 
                                    "phase", 
                                    "maxvisit",
                                    "age_bl", 
                                    "sex_bl",
                                    "height_bl",
                                    "weight_bl",
                                    "race_bl",
                                    "ethnic_bl",
                                    "dtvisit", 
                                    "intervention",
                                    "antibiotic_bl", 
                                    "feedtube", 
                                    "urincath", 
                                    "deviceuse", 
                                    "openwound_final",
                                    "cogsum_bl",
                                    "homrsa_bl",
                                    "hovre_bl",
                                    "horgnb_bl",
                                    "psmscore",
                                    "anyanypos_p",
                                    "anymrsa_p",
                                    "anyvre_p",
                                    "anygnbr_p",
                                    "acqmrsa", 
                                    "acqvre", 
                                    "acqgnbr", 
                                    "hospstay", 
                                    "charlson", 
                                    "ccsmi_bl", 
                                    "ccschf_bl", 
                                    "ccspvod_bl", 
                                    "ccscva_bl", 
                                    "ccsdemen_bl", 
                                    "ccscopd_bl", 
                                    "ccscontis_bl", 
                                    "ccspud_bl", 
                                    "ccsmld_bl", 
                                    "ccsdm_bl", 
                                    "ccshemi_bl", 
                                    "ccsmsrd_bl", 
                                    "ccsdmeod_bl", 
                                    "ccstumor_bl", 
                                    "ccsleuk_bl", 
                                    "ccslymph_bl", 
                                    "ccsmsld_bl",
                                   "ccsmst_bl",
                                    "ccsaids_bl",
                                "comorbidity_category",
                                    "VREsiteE1",
                                    "VREsiteE10",
                                    "VREsiteE11",
                                    "VREsiteE12",
                                    "VREsiteE2",
                                    "VREsiteE3", 
                                    "VREsiteE4",
                                    "VREsiteE5",
                                    "VREsiteE6",
                                    "VREsiteE7",
                                    "VREsiteE8",
                                    "VREsiteE9",
                                    "VREsiteF",
                                    "VREsiteG",
                                    "VREsiteH",
                                    "VREsiteN",
                                    "VREsiteO",
                                    "VREsiteR",
                                    "VREsiteS",
                                    "vre_p",
                                    "BMI",
                                   "visit_intervened",
                                    "gender",
                                    "race",
                                    "charlson_cat",
                                "vre_env",
                                "vre_p_cat",
                                "vre_env_cat"
                                    )]
```

## Colonization Subsets with different MDROs
```{r}
# Select the columns you want to keep from rgnb_colonization_persistence
rgnb_cols <- rgnb_colonization_persistence[, c("studyID", "persistence", "cleared", "recnum", "duration")]

# Merge test_subset and rgnb_cols by studyID
rgnb_c <- merge(test_subset_rgnb, rgnb_cols, by = "studyID", all.x = TRUE)

# Keep only observations where "persistence" is not NA
rgnb_c <- rgnb_c[!is.na(rgnb_c$persistence), ]

# Copy labels from test_subset to rgnb_c
rgnb_c <- copy_labels(test_subset, rgnb_c)

# Set a label for multiple sites
set_label(rgnb_c$gnbr_p_cat) <- "Number of patient sites colonization"

# Select the columns you want to keep from mrsa_colonization_persistence
mrsa_cols <- mrsa_colonization_persistence[, c("studyID", "persistence", "cleared", "recnum", "duration")]

# Merge test_subset and mrsa_cols by studyID
mrsa_c <- merge(test_subset_mrsa, mrsa_cols, by = "studyID", all.x = TRUE)

# Keep only observations where "persistence" is not NA
mrsa_c <- mrsa_c[!is.na(mrsa_c$persistence), ]

# Copy labels from test_subset to mrsa_c
mrsa_c <- copy_labels(test_subset, mrsa_c)

# Set a label for multiple sites
set_label(mrsa_c$mrsa_p_cat) <- "Number of patient sites colonization"

# Select the columns you want to keep from vre_colonization_persistence
vre_cols <- vre_colonization_persistence[, c("studyID", "persistence", "cleared", "recnum", "duration")]

# Merge test_subset and vre_cols by studyID
vre_c <- merge(test_subset_vre, vre_cols, by = "studyID", all.x = TRUE)

# Keep only observations where "persistence" is not NA
vre_c <- vre_c[!is.na(vre_c$persistence), ]

# Copy labels from test_subset to vre_c
vre_c <- copy_labels(test_subset, vre_c)

# Set a label for multiple sites
set_label(vre_c$vre_p_cat) <- "Number of patient sites colonization"

# Remove duplicate study IDs
rgnb_unique_c <- rgnb_c %>%
  filter(visit == 0) %>%
  distinct(studyID, .keep_all = TRUE)

# Replace persistence values with labels
rgnb_unique_c <- rgnb_unique_c %>%
mutate(persistence = factor(persistence, levels = c(0, 1), labels = c("Intermittent", "Persistent")))%>%
  mutate(cleared = factor(cleared, levels = c(0, 1), labels = c("No", "Yes")))
set_label(rgnb_unique_c$persistence) <- "Persistence"
set_label(rgnb_unique_c$cleared) <- "Clearance"

# Remove duplicate study IDs
mrsa_unique_c <- mrsa_c %>%
    filter(visit == 0) %>%
  distinct(studyID, .keep_all = TRUE)

# Replace persistence values with labels
mrsa_unique_c <- mrsa_unique_c %>%
mutate(persistence = factor(persistence, levels = c(0, 1), labels = c("Intermittent", "Persistent")))%>%
  mutate(cleared = factor(cleared, levels = c(0, 1), labels = c("No", "Yes")))
set_label(mrsa_unique_c$persistence) <- "Persistence"
set_label(mrsa_unique_c$cleared) <- "Clearance"

# Remove duplicate study IDs
vre_unique_c <- vre_c %>%
    filter(visit == 0) %>%
  distinct(studyID, .keep_all = TRUE)

# Replace persistence values with labels
vre_unique_c <- vre_unique_c %>%
mutate(persistence = factor(persistence, levels = c(0, 1), labels = c("Intermittent", "Persistent"))) %>%
  mutate(cleared = factor(cleared, levels = c(0, 1), labels = c("No", "Yes")))
set_label(vre_unique_c$persistence) <- "Persistence"
set_label(vre_unique_c$cleared) <- "Clearance"
```

## New Acquisition Subsets with different MDROs
```{r}
# Select the columns you want to keep from rgnb_acquisition_persistence
rgnb_cols <- rgnb_acquisition_persistence[, c("studyID", "persistence", "cleared", "recnum", "duration")]

# Merge test_subset and rgnb_cols by studyID
rgnb_a <- merge(test_subset, rgnb_cols, by = "studyID", all.x = TRUE)

# Keep only observations where "persistence" is not NA
rgnb_a <- rgnb_a[!is.na(rgnb_a$persistence), ]

# Copy labels from test_subset to rgnb_a
rgnb_a <- copy_labels(test_subset, rgnb_a)

# Set a label for multiple sites
set_label(rgnb_a$gnbr_p_cat) <- "Number of patient sites colonization"

# Select the columns you want to keep from mrsa_acquisition_persistence
mrsa_cols <- mrsa_acquisition_persistence[, c("studyID", "persistence", "cleared", "recnum", "duration")]

# Merge test_subset and mrsa_cols by studyID
mrsa_a <- merge(test_subset, mrsa_cols, by = "studyID", all.x = TRUE)

# Keep only observations where "persistence" is not NA
mrsa_a <- mrsa_a[!is.na(mrsa_a$persistence), ]

# Copy labels from test_subset to mrsa_a
mrsa_a <- copy_labels(test_subset, mrsa_a)

# Set a label for multiple sites
set_label(mrsa_a$mrsa_p_cat) <- "Number of patient sites colonization"

# Select the columns you want to keep from vre_acquisition_persistence
vre_cols <- vre_acquisition_persistence[, c("studyID", "persistence", "cleared", "recnum","duration")]

# Merge test_subset and vre_cols by studyID
vre_a <- merge(test_subset, vre_cols, by = "studyID", all.x = TRUE)

# Keep only observations where "persistence" is not NA
vre_a <- vre_a[!is.na(vre_a$persistence), ]

# Copy labels from test_subset to vre_a
vre_a <- copy_labels(test_subset, vre_a)

# Set a label for multiple sites
set_label(vre_a$vre_p_cat) <- "Number of patient sites colonization"

# Remove duplicate study IDs, keeping the one where acqgnbr = 1
rgnb_unique_a <- rgnb_a %>%
  group_by(studyID) %>%
  arrange(desc(acqgnbr)) %>%  # Sort in descending order, so 1 comes before 0
  slice(1) %>%  # Keep only the first row (which will be acqgnrb = 1 if it exists)
  ungroup()

# Replace persistence values with labels
rgnb_unique_a <- rgnb_unique_a %>%
mutate(persistence = factor(persistence, levels = c(0, 1), labels = c("Intermittent", "Persistent")))%>%
  mutate(cleared = factor(cleared, levels = c(0, 1), labels = c("No", "Yes")))
set_label(rgnb_unique_a$persistence) <- "Persistence"
set_label(rgnb_unique_a$cleared) <- "Clearance"

# Remove duplicate study IDs, keeping the one where acqmrsa = 1
mrsa_unique_a <- mrsa_a %>%
  group_by(studyID) %>%
  arrange(desc(acqmrsa)) %>%  # Sort in descending order, so 1 comes before 0
  slice(1) %>%  # Keep only the first row (which will be acqgnrb = 1 if it exists)
  ungroup()

# Replace persistence values with labels
mrsa_unique_a <- mrsa_unique_a %>%
mutate(persistence = factor(persistence, levels = c(0, 1), labels = c("Intermittent", "Persistent")))%>%
  mutate(cleared = factor(cleared, levels = c(0, 1), labels = c("No", "Yes")))
set_label(mrsa_unique_a$persistence) <- "Persistence"
set_label(mrsa_unique_a$cleared) <- "Clearance"

# Remove duplicate study IDs, keeping the one where acqvre = 1
vre_unique_a <- vre_a %>%
  group_by(studyID) %>%
  arrange(desc(acqvre)) %>%  # Sort in descending order, so 1 comes before 0
  slice(1) %>%  # Keep only the first row (which will be acqgnrb = 1 if it exists)
  ungroup()

# Replace persistence values with labels
vre_unique_a <- vre_unique_a %>%
mutate(persistence = factor(persistence, levels = c(0, 1), labels = c("Intermittent", "Persistent"))) %>%
  mutate(cleared = factor(cleared, levels = c(0, 1), labels = c("No", "Yes")))
set_label(vre_unique_a$persistence) <- "Persistence"
set_label(vre_unique_a$cleared) <- "Clearance"

# Combine Unique Table
combined_unique_a <- bind_rows(rgnb_unique_a, mrsa_unique_a, vre_unique_a, .id = "organism")
combined_unique_a <- combined_unique_a %>%
  mutate(organism = case_when(
    organism == 1 ~ "RGNB",
    organism == 2 ~ "MRSA",
    organism == 3 ~ "VRE"
  ))

combined_unique_c <- bind_rows(rgnb_unique_c, mrsa_unique_c, vre_unique_c, .id = "organism")
combined_unique_c <- combined_unique_c %>%
  mutate(organism = case_when(
    organism == 1 ~ "RGNB",
    organism == 2 ~ "MRSA",
    organism == 3 ~ "VRE"
  ))

# Create New Variable (New Acquisition Season)
```

# Export datasets
```{r}
# Export rgnb_a to CSV file
write.csv(rgnb_a, file = "rgnb_acquisition_complete.csv", row.names = FALSE)

# Export rgnb_c to CSV file
write.csv(rgnb_c, file = "rgnb_colonization_complete.csv", row.names = FALSE)

# Export mrsa_a to CSV file
write.csv(mrsa_a, file = "mrsa_acquisition_complete.csv", row.names = FALSE)

# Export mrsa_c to CSV file
write.csv(mrsa_c, file = "mrsa_colonization_complete.csv", row.names = FALSE)

# Export vre_a to CSV file
write.csv(vre_a, file = "vre_acquisition_complete.csv", row.names = FALSE)

# Export vre_c to CSV file
write.csv(vre_c, file = "vre_colonization_complete.csv", row.names = FALSE)
```

# Table 1
## Descriptive Statistics for each MDROs (Colonization)
```{r}
# Summary for persistence in rgnb_unique
summary_persistence_rgnb_c <- rgnb_unique_c %>%
  dplyr::select(persistence, cleared) %>%
  tbl_summary(
    type = list(persistence ~ "categorical",
                cleared ~ "categorical"),
    statistic = all_categorical() ~ "{n} ({p}%)",
    digits = list(all_categorical() ~ c(0, 1)),
    missing = "no"
    #missing_text = "Missing (n)"
  ) %>%
  modify_header(label ~ "**Variable**") %>%
  bold_labels() %>%
  modify_spanning_header(all_stat_cols() ~ "**RGNB**")

# Summary for persistence in mrsa_unique
summary_persistence_mrsa_c <- mrsa_unique_c %>%
  dplyr::select(persistence, cleared) %>%
  tbl_summary(
    type = list(persistence ~ "categorical",
                cleared ~ "categorical"),
    statistic = all_categorical() ~ "{n} ({p}%)",
    digits = list(all_categorical() ~ c(0, 1)),
    missing = "no"
  ) %>%
  modify_header(label ~ "**Variable**") %>%
  bold_labels() %>%
  modify_spanning_header(all_stat_cols() ~ "**MRSA**")

# Summary for persistence in vre_unique
summary_persistence_vre_c <- vre_unique_c %>%
  dplyr::select(persistence, cleared) %>%
  tbl_summary(
    type = list(persistence ~ "categorical",
                cleared ~ "categorical"),
    statistic = all_categorical() ~ "{n} ({p}%)",
    digits = list(all_categorical() ~ c(0, 1)),
    missing = "no"
  ) %>%
  modify_header(label ~ "**Variable**") %>%
  bold_labels() %>%
  modify_spanning_header(all_stat_cols() ~ "**VRE**")

# Combine summaries for persistence
combined_persistence_summary_c <- tbl_merge(
  tbls = list(summary_persistence_rgnb_c, summary_persistence_mrsa_c, summary_persistence_vre_c),
  tab_spanner = c("**RGNB**", "**MRSA**", "**VRE**")
) %>%
as_gt() %>%
tab_header(
  title = "Persistence and Clearance among NH Residents with Preexisting Colonization"
) %>%
tab_style(
  cell_borders(sides = c("top"), style = "hidden"),
  locations = cells_title()
)

# View and export the combined persistence table
combined_persistence_summary_c
```
## Add chi-square p-value in persistence among organisms
```{r}
# Define the contingency tables
table1 <- matrix(c(37, 14, 30, 81, 54, 104), nrow = 2, byrow = TRUE)
table2 <- matrix(c(70, 47, 85, 11, 7, 19), nrow = 2, byrow = TRUE)
table3 <- matrix(c(69, 28, 50, 47, 27, 29), nrow = 2, byrow = TRUE)
table4 <- matrix(c(37, 26, 25, 10, 1, 4), nrow = 2, byrow = TRUE)

# Function to decide which test to use and perform it
perform_test <- function(table) {
  # Calculate expected frequencies
  expected_frequencies <- chisq.test(table)$expected
  
  # Decide which test to use
  if (all(expected_frequencies >= 5)) {
    # Use chi-square test
    test_result <- chisq.test(table)
  } else {
    # Use Fisher's exact test
    test_result <- fisher.test(table)
  }
  
  return(test_result)
}

# Perform tests on all tables
test_result1 <- perform_test(table1)
test_result2 <- perform_test(table2)
test_result3 <- perform_test(table3)
test_result4 <- perform_test(table4)

# Print results
table1
print(test_result1)
table2
print(test_result2)
table3
print(test_result3)
table4
print(test_result4)
```


## Descriptive Statistics for each MDROs (New_acquisition)
```{r}
# Summary for persistence in rgnb_unique
summary_persistence_rgnb_a <- rgnb_unique_a %>%
  dplyr::select(persistence, cleared) %>%
  tbl_summary(
    type = list(persistence ~ "categorical",
                cleared ~ "categorical"),
    statistic = all_categorical() ~ "{n} ({p}%)",
    digits = list(all_categorical() ~ c(0, 1)),
    missing = "no"
    #missing_text = "Missing (n)"
  ) %>%
  modify_header(label ~ "**Variable**") %>%
  bold_labels() %>%
  modify_spanning_header(all_stat_cols() ~ "**RGNB**")

# Summary for persistence in mrsa_unique
summary_persistence_mrsa_a <- mrsa_unique_a %>%
  dplyr::select(persistence, cleared) %>%
  tbl_summary(
    type = list(persistence ~ "categorical",
                cleared ~ "categorical"),
    statistic = all_categorical() ~ "{n} ({p}%)",
    digits = list(all_categorical() ~ c(0, 1)),
    missing = "no"
  ) %>%
  modify_header(label ~ "**Variable**") %>%
  bold_labels() %>%
  modify_spanning_header(all_stat_cols() ~ "**MRSA**")

# Summary for persistence in vre_unique
summary_persistence_vre_a <- vre_unique_a %>%
  dplyr::select(persistence, cleared) %>%
  tbl_summary(
    type = list(persistence ~ "categorical",
                cleared ~ "categorical"),
    statistic = all_categorical() ~ "{n} ({p}%)",
    digits = list(all_categorical() ~ c(0, 1)),
    missing = "no"
  ) %>%
  modify_header(label ~ "**Variable**") %>%
  bold_labels() %>%
  modify_spanning_header(all_stat_cols() ~ "**VRE**")

# Combine summaries for persistence
combined_persistence_summary_a <- tbl_merge(
  tbls = list(summary_persistence_rgnb_a, summary_persistence_mrsa_a, summary_persistence_vre_a),
  tab_spanner = c("**RGNB**", "**MRSA**", "**VRE**")
) %>%
as_gt() %>%
tab_header(
  title = "Persistence and Clearance among NH Residents with New Acquisition"
) %>%
tab_style(
  cell_borders(sides = c("top"), style = "hidden"),
  locations = cells_title()
)

# View and export the combined persistence table
combined_persistence_summary_a
```

# Table 2
## Duration (Colonization) with persistence
```{r}
# Function to summarize duration (with persistence)
summarize_duration <- function(data, label) {
  data %>%
    filter(persistence == "Persistent") %>%
    summarize(
      N = n(),
      Mean = mean(duration, na.rm = TRUE),
      SD = sd(duration, na.rm = TRUE),
      Median = median(duration, na.rm = TRUE),
      Min = min(duration, na.rm = TRUE),
      Max = max(duration, na.rm = TRUE)
    ) %>%
    mutate(Variable = label) %>%
    select(Variable, everything())
}

# Function to summarize duration (cleared)
summarize_duration_c <- function(data, label) {
  data %>%
    filter(cleared == "Yes") %>%
    summarize(
      N = n(),
      Mean = mean(duration, na.rm = TRUE),
      SD = sd(duration, na.rm = TRUE),
      Median = median(duration, na.rm = TRUE),
      Min = min(duration, na.rm = TRUE),
      Max = max(duration, na.rm = TRUE)
    ) %>%
    mutate(Variable = label) %>%
    select(Variable, everything())
}
```

```{r}
# Summarize duration for each dataset
summary_duration_rgnb_c <- summarize_duration(rgnb_unique_c, "RGNB")
summary_duration_mrsa_c <- summarize_duration(mrsa_unique_c, "MRSA")
summary_duration_vre_c <- summarize_duration(vre_unique_c, "VRE")

# Combine the summaries into one dataframe
combined_duration_summary_c <- bind_rows(summary_duration_rgnb_c, summary_duration_mrsa_c, summary_duration_vre_c)

# Convert to gt table
duration_gt_table_c <- combined_duration_summary_c %>%
  gt() %>%
  tab_header(
    title = "Duration of Persistence in Preexisting Colonization - Persistent Patients"
  ) %>%
  fmt_number(
    columns = c(Mean, SD, Median, Min, Max),
    decimals = 2
  ) %>%
    tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_body(columns = Variable)
  ) %>%
  tab_options(
    column_labels.font.weight = "bold"
  )
  
# View the combined duration table
duration_gt_table_c
```
## Duration (Colonization) with cleared
```{r}
# Summarize duration for each dataset
summary_duration_rgnb_c_c <- summarize_duration_c(rgnb_unique_c, "RGNB")
summary_duration_mrsa_c_c <- summarize_duration_c(mrsa_unique_c, "MRSA")
summary_duration_vre_c_c <- summarize_duration_c(vre_unique_c, "VRE")

# Combine the summaries into one dataframe
combined_duration_summary_c_c <- bind_rows(summary_duration_rgnb_c_c, summary_duration_mrsa_c_c, summary_duration_vre_c_c)

# Convert to gt table
duration_gt_table_c_c <- combined_duration_summary_c_c %>%
  gt() %>%
  tab_header(
    title = "Duration of Persistence in Preexisting Colonization - Cleared Patients"
  ) %>%
  fmt_number(
    columns = c(Mean, SD, Median, Min, Max),
    decimals = 2
  ) %>%
    tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_body(columns = Variable)
  ) %>%
  tab_options(
    column_labels.font.weight = "bold"
  )

# View the combined duration table
duration_gt_table_c_c
```
## Duration (NewAcquisition) with persistence
```{r}
# Summarize duration for each dataset
summary_duration_rgnb_a <- summarize_duration(rgnb_unique_a, "RGNB")
summary_duration_mrsa_a <- summarize_duration(mrsa_unique_a, "MRSA")
summary_duration_vre_a <- summarize_duration(vre_unique_a, "VRE")

# Combine the summaries into one dataframe
combined_duration_summary_a <- bind_rows(summary_duration_rgnb_a, summary_duration_mrsa_a, summary_duration_vre_a)

# Convert to gt table
duration_gt_table_a <- combined_duration_summary_a %>%
  gt() %>%
  tab_header(
    title = "Duration of Persistence in New Acquisition - Persistent Patients"
  ) %>%
  fmt_number(
    columns = c(Mean, SD, Median, Min, Max),
    decimals = 2
  ) %>%
    tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_body(columns = Variable)
  ) %>%
  tab_options(
    column_labels.font.weight = "bold"
  )
  
# View the combined duration table
duration_gt_table_a
```
## Duration (NewAcquisition) with cleared
```{r}
# Summarize duration for each dataset
summary_duration_rgnb_a_c <- summarize_duration_c(rgnb_unique_a, "RGNB")
summary_duration_mrsa_a_c <- summarize_duration_c(mrsa_unique_a, "MRSA")
summary_duration_vre_a_c <- summarize_duration_c(vre_unique_a, "VRE")

# Combine the summaries into one dataframe
combined_duration_summary_a_c <- bind_rows(summary_duration_rgnb_a_c, summary_duration_mrsa_a_c, summary_duration_vre_a_c)

# Convert to gt table
duration_gt_table_a_c <- combined_duration_summary_a_c %>%
  gt() %>%
  tab_header(
    title = "Duration of Persistence in New Acquisition - Cleared Patients"
  ) %>%
  fmt_number(
    columns = c(Mean, SD, Median, Min, Max),
    decimals = 2
  ) %>%
    tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_body(columns = Variable)
  ) %>%
  tab_options(
    column_labels.font.weight = "bold"
  )

# View the combined duration table
duration_gt_table_a_c
```
# Figure 1.Compare duration in perexisting vs. new acquisition
```{r}
# Combine data frames for pre-existing and new acquisition
combined_duration_rgnb <- bind_rows(
  rgnb_unique_c %>% mutate(Colonization = "Pre-existing"),
  rgnb_unique_a %>% mutate(Colonization = "New Acquisition"),
  .id = "id")

# Combine data frames for pre-existing and new acquisition
combined_duration_mrsa <- bind_rows(
  mrsa_unique_c %>% mutate(Colonization = "Pre-existing"),
 mrsa_unique_a %>% mutate(Colonization = "New Acquisition"),
  .id = "id")

# Combine data frames for pre-existing and new acquisition
combined_duration_vre <- bind_rows(
  vre_unique_c %>% mutate(Colonization = "Pre-existing"),
  vre_unique_a %>% mutate(Colonization = "New Acquisition"),
  .id = "id")

# Combine all of them
combined_duration <- bind_rows(combined_duration_rgnb, combined_duration_mrsa, combined_duration_vre, .id = "organism")
combined_duration <- combined_duration %>%
  mutate(organism = case_when(
    organism == 1 ~ "RGNB",
    organism == 2 ~ "MRSA",
    organism == 3 ~ "VRE"
  ))

# Check the distribution
hist(combined_duration_rgnb$duration, main="Histogram of Duration"
, xlab="Duration")
qqnorm(combined_duration_rgnb$duration, main = "Q-Q Plot of Duration")
qqline(combined_duration_rgnb$duration) # Not normally distribution
shapiro.test(combined_duration_rgnb$duration)

hist(combined_duration_mrsa$duration, main="Histogram of Duration"
, xlab="Duration")
qqnorm(combined_duration_mrsa$duration, main = "Q-Q Plot of Duration")
qqline(combined_duration_mrsa$duration) # Not normally distribution
shapiro.test(combined_duration_mrsa$duration)

hist(combined_duration_vre$duration, main="Histogram of Duration"
, xlab="Duration")
qqnorm(combined_duration_vre$duration, main = "Q-Q Plot of Duration")
qqline(combined_duration_vre$duration) # Not normally distribution
shapiro.test(combined_duration_vre$duration)

# Create boxplot with p values
p1 <- ggboxplot(combined_duration, 
                x = "Colonization", y = "duration",
                color = "Colonization", palette = "jco",
                add = "jitter",
                facet.by = "organism") +
  stat_compare_means(method = "wilcox.test") + 
  labs(y = "Duration of persistence")  # Add y-axis label
p1

ggsave("boxplot_1.png", plot = p1, width = 8, height = 6, dpi = 600)
```

```{r}
library(dplyr)
library(ggpubr)
library(ggplot2)

# Filter data where cleared = "Yes"
rgnb_unique_c_yes <- rgnb_unique_c %>% filter(cleared == "Yes")
rgnb_unique_a_yes <- rgnb_unique_a %>% filter(cleared == "Yes")
mrsa_unique_c_yes <- mrsa_unique_c %>% filter(cleared == "Yes")
mrsa_unique_a_yes <- mrsa_unique_a %>% filter(cleared == "Yes")
vre_unique_c_yes <- vre_unique_c %>% filter(cleared == "Yes")
vre_unique_a_yes <- vre_unique_a %>% filter(cleared == "Yes")

# Combine data frames for pre-existing and new acquisition for RGNB
combined_duration_rgnb_2 <- bind_rows(
  rgnb_unique_c_yes %>% mutate(Colonization = "Pre-existing"),
  rgnb_unique_a_yes %>% mutate(Colonization = "New Acquisition"),
  .id = "id"
)

# Combine data frames for pre-existing and new acquisition for MRSA
combined_duration_mrsa_2 <- bind_rows(
  mrsa_unique_c_yes %>% mutate(Colonization = "Pre-existing"),
  mrsa_unique_a_yes %>% mutate(Colonization = "New Acquisition"),
  .id = "id"
)

# Combine data frames for pre-existing and new acquisition for VRE
combined_duration_vre_2 <- bind_rows(
  vre_unique_c_yes %>% mutate(Colonization = "Pre-existing"),
  vre_unique_a_yes %>% mutate(Colonization = "New Acquisition"),
  .id = "id"
)

# Combine all of them
combined_duration_2 <- bind_rows(combined_duration_rgnb_2, combined_duration_mrsa_2, combined_duration_vre_2, .id = "organism")
combined_duration_2 <- combined_duration_2 %>%
  mutate(organism = case_when(
    organism == 1 ~ "RGNB",
    organism == 2 ~ "MRSA",
    organism == 3 ~ "VRE"
  ))

# Create boxplot with p values
p1_2 <- ggboxplot(combined_duration_2, 
                  x = "Colonization", y = "duration",
                  color = "Colonization", palette = "jco",
                  add = "jitter",
                  facet.by = "organism") +
  stat_compare_means(method = "wilcox.test") + 
  labs(y = "Duration of persistence")  # Add y-axis label

# Display the plot
print(p1_2)

# Save the plot as a PNG file
ggsave("boxplot_2.png", plot = p1_2, width = 8, height = 6, dpi = 600)
```
# Figure 2. Compare duration with clearance of pre-existing colonization among species
```{r}
library(dplyr)
library(ggpubr)
library(ggplot2)

# Perorm pairwise comparisons
compare_means(duration ~ organism,  data = combined_duration_2)

# Change Colonization Order
combined_duration_2$Colonization = factor(combined_duration_2$Colonization,
                                          levels = c("Pre-existing", "New Acquisition"))

# Create boxplot with p values
my_comparisons <- list( c("RGNB", "MRSA"), c("MRSA", "VRE"), c("RGNB", "VRE") )
p2 <- ggboxplot(combined_duration_2, 
                x = "organism", 
                y = "duration",
                color = "organism", 
                palette = "jco",
                add = "jitter",
                facet.by = "Colonization") +
  stat_compare_means(comparisons = my_comparisons)+ # Add pairwise comparisons p-value
  stat_compare_means(label.y = 100)     # Add global p-value + 
  labs(y = "Duration of persistence")  # Add y-axis label

# Display the plot
print(p2)

# Save the plot as a PNG file
ggsave("boxplot_3.png", plot = p2, width = 8, height = 6, dpi = 600)
```



# Table 3: Univariate Analysis for Persistence Outcome
## RGNB Univariate_Colonization
### Check Distribution
```{r}

```

### Generate table
```{r}
#| label: RGNB univariate analysis_colonization
# Colonization RGNB Univariate Table (Crude)
rgnb_uni_c <- rgnb_unique_c %>%
  dplyr::select(age_bl,
                BMI,
                gender,
                race,
                gnbr_p_cat,
                gnbr_env_cat,
                # ethnic_bl,# Only have 2 = hispanic or latino
                antibiotic_bl,
                deviceuse, 
                feedtube,
                urincath, 
                openwound_final,
                cogsum_bl,
                horgnb_bl,
                psmscore,
                charlson_cat,
                hospstay,
                visit_intervened,
                persistence,
                ccsmi_bl, 
                ccschf_bl, 
                ccspvod_bl, 
                ccscva_bl, 
                ccsdemen_bl, 
                ccscopd_bl, 
                ccscontis_bl, 
                ccspud_bl, 
                ccsmld_bl, 
                ccsdm_bl, 
                ccshemi_bl, 
                ccsmsrd_bl, 
                ccsdmeod_bl, 
                ccstumor_bl, 
                # ccsleuk_bl, # only have 0
                # ccslymph_bl, # only have 0
                ccsmsld_bl, 
                ccsmst_bl, 
                # ccsaids_bl # only have 0
) %>%
  tbl_summary(by = persistence, # stratify by persistence
              statistic = list(all_continuous() ~ "{mean} ({sd})", #use mean and standard deviation for continuous variables
                               all_categorical() ~ "{n} ({p}%)"), #use count and percentage for categorical variables
              digits = list(all_categorical() ~ c(0, 2), #adds no decimal places to counts and one decimal to percentages
                            all_continuous() ~ 1), #adds one decimal place to mean and sd values
              missing = "no"
              #missing_text = "Missing (n)"
              ) %>%
  add_p(pvalue_fun = ~style_pvalue(.x, digits = 2)) %>% #adds p value
  add_overall() %>% #adds column with non-stratified summary statistics
  modify_header(label ~ "**Variable**") %>% #the asterisks bold the label
  modify_spanning_header(c("stat_1", "stat_2") ~ "**Persistence**") %>%
  bold_labels() %>%
  as_gt() %>% 
  tab_header(
    title = "Table 3. Variables associated with RGNB persistence in perexisting colonization patient"
    ) %>% 
  tab_style(cell_borders(
    sides = c("top"), style = "hidden"),
    locations = cells_title())

# View the table
rgnb_uni_c
```
## MRSA Univariate_Colonization
```{r}
#| label: MRSA univariate analysis_colonization
# Colonization MRSA Univariate Table (Crude)
mrsa_uni_c <- mrsa_unique_c %>%
  dplyr::select(age_bl,
                BMI,
                gender,
                race,
                mrsa_p_cat,
                mrsa_env_cat,
                # ethnic_bl,# Only have 2 = hispanic or latino
                antibiotic_bl,
                deviceuse, 
                feedtube,
                urincath, 
                openwound_final,
                cogsum_bl,
                homrsa_bl,
                psmscore,
                charlson_cat,
                hospstay,
                visit_intervened,
                persistence,
                ccsmi_bl, 
                ccschf_bl, 
                ccspvod_bl, 
                ccscva_bl, 
                ccsdemen_bl, 
                ccscopd_bl, 
                ccscontis_bl, 
                ccspud_bl, 
                ccsmld_bl, 
                ccsdm_bl, 
                ccshemi_bl, 
                ccsmsrd_bl, 
                ccsdmeod_bl, 
                ccstumor_bl, 
                # ccsleuk_bl, # only have 0
                # ccslymph_bl,# only have 0
                ccsmsld_bl, 
                ccsmst_bl, 
                # ccsaids_bl # only have 0
                ) %>%
  tbl_summary(by = persistence, # stratify by outcome
              statistic = list(all_continuous() ~ "{mean} ({sd})", #use mean and standard deviation for continuous variables
                               all_categorical() ~ "{n} ({p}%)"), #use count and percentage for categorical variables
              digits = list(all_categorical() ~ c(0, 2), #adds no decimal places to counts and one decimal to percentages
                            all_continuous() ~ 1), #adds one decimal place to mean and sd values
              missing = "no"
              #missing_text = "Missing (n)"
              ) %>%
  add_p(pvalue_fun = ~style_pvalue(.x, digits = 2)) %>% #adds p value
  add_overall() %>% #adds column with non-stratified summary statistics
  modify_header(label ~ "**Variable**") %>% #the asterisks bold the label
  modify_spanning_header(c("stat_1", "stat_2") ~ "**Persistence**") %>%
  bold_labels() %>%
  as_gt() %>% 
  tab_header(
    title = "Table 3. Variables associated with MRSA persistence in perexisting colonization patient"
    ) %>% 
  tab_style(cell_borders(
    sides = c("top"), style = "hidden"),
    locations = cells_title())

# View the table
mrsa_uni_c
```
## VRE Univariate_Colonization
```{r}
#| label: VRE univariate analysis_colonization
# Colonization VRE Univariate Table (Crude)
vre_uni_c <- vre_unique_c %>%
  dplyr::select(age_bl,
                BMI,
                gender,
                race,
                vre_p_cat,
                vre_env_cat,
                # ethnic_bl,# 1 only have 1 observation
                antibiotic_bl,
                deviceuse, 
                feedtube,
                urincath, 
                openwound_final,
                cogsum_bl,
                hovre_bl,
                psmscore,
                charlson_cat,
                hospstay,
                visit_intervened,
                persistence,
                ccsmi_bl, 
                ccschf_bl, 
                ccspvod_bl, 
                ccscva_bl, 
                ccsdemen_bl, 
                ccscopd_bl, 
                ccscontis_bl, 
                ccspud_bl, 
                ccsmld_bl, 
                ccsdm_bl, 
                ccshemi_bl, 
                ccsmsrd_bl, 
                ccsdmeod_bl, 
                ccstumor_bl, 
                # ccsleuk_bl, # only have 0
                # ccslymph_bl,# 1 only have 1 observation
                ccsmsld_bl, 
                ccsmst_bl, 
                # ccsaids_bl # only have 0
                ) %>%
  tbl_summary(by = persistence, # stratify by outcome
              statistic = list(all_continuous() ~ "{mean} ({sd})", #use mean and standard deviation for continuous variables
                               all_categorical() ~ "{n} ({p}%)"), #use count and percentage for categorical variables
              digits = list(all_categorical() ~ c(0, 2), #adds no decimal places to counts and one decimal to percentages
                            all_continuous() ~ 1), #adds one decimal place to mean and sd values
              missing = "no"
              #missing_text = "Missing (n)"
              ) %>%
  add_p(pvalue_fun = ~style_pvalue(.x, digits = 2)) %>% #adds p value
  add_overall() %>% #adds column with non-stratified summary statistics
  modify_header(label ~ "**Variable**") %>% #the asterisks bold the label
  modify_spanning_header(c("stat_1", "stat_2") ~ "**Persistence**") %>%
  bold_labels() %>%
  as_gt() %>% 
  tab_header(
    title = "Table 3. Variables associated with VRE persistence in perexisting colonization patient"
    ) %>% 
  tab_style(cell_borders(
    sides = c("top"), style = "hidden"),
    locations = cells_title())

# View the table
vre_uni_c
```

## RGNB Univariate_Acquisition
```{r}
#| label: RGNB univariate analysis_acquisition
# Colonization RGNB Univariate Table (Crude)
rgnb_uni_a <- rgnb_unique_a %>%
  dplyr::select(age_bl,
                BMI,
                gender,
                race,
                gnbr_p_cat,
                gnbr_env_cat,
                # ethnic_bl,# Only have 1 observation in 1
                antibiotic_bl,
                deviceuse, 
                feedtube,
                urincath, 
                openwound_final,
                cogsum_bl,
                horgnb_bl,
                psmscore,
                charlson_cat,
                hospstay,
                visit_intervened,
                persistence,
                ccsmi_bl, 
                ccschf_bl, 
                ccspvod_bl, 
                ccscva_bl, 
                ccsdemen_bl, 
                ccscopd_bl, 
                ccscontis_bl, 
                ccspud_bl, 
                ccsmld_bl, 
                ccsdm_bl, 
                ccshemi_bl, 
                ccsmsrd_bl, 
                ccsdmeod_bl, 
                ccstumor_bl, 
                ccsleuk_bl,
                ccslymph_bl,
                # ccsmsld_bl, # only have 0
                ccsmst_bl,
                # ccsaids_bl, # NA
                ) %>%
  tbl_summary(by = persistence, # stratify by outcome
              statistic = list(all_continuous() ~ "{mean} ({sd})", #use mean and standard deviation for continuous variables
                               all_categorical() ~ "{n} ({p}%)"), #use count and percentage for categorical variables
              digits = list(all_categorical() ~ c(0, 2), #adds no decimal places to counts and one decimal to percentages
                            all_continuous() ~ 1), #adds one decimal place to mean and sd values
              missing = "no"
              #missing_text = "Missing (n)"
              ) %>%
  add_p(pvalue_fun = ~style_pvalue(.x, digits = 2)) %>% #adds p value
  add_overall() %>% #adds column with non-stratified summary statistics
  modify_header(label ~ "**Variable**") %>% #the asterisks bold the label
  modify_spanning_header(c("stat_1", "stat_2") ~ "**Persistence**") %>%
  bold_labels() %>%
  as_gt() %>% 
  tab_header(
    title = "Table 4. Variables associated with RGNB persistence in new acquisition patient"
    ) %>% 
  tab_style(cell_borders(
    sides = c("top"), style = "hidden"),
    locations = cells_title())

# View the table
rgnb_uni_a
```
## MRSA Univariate_Acquisition
```{r}
#| label: MRSA univariate analysis_acquisition
# Colonization MRSA Univariate Table (Crude)
mrsa_uni_a <- mrsa_unique_a %>%
  dplyr::select(age_bl,
                BMI,
                gender,
                race,
                mrsa_p_cat,
                mrsa_env_cat,
                # ethnic_bl,# Only have 1 observation in 1
                antibiotic_bl,
                deviceuse, 
                feedtube,
                urincath, 
                openwound_final,
                cogsum_bl,
                homrsa_bl,
                psmscore,
                charlson_cat,
                hospstay,
                visit_intervened,
                persistence,
                ccsmi_bl, 
                ccschf_bl, 
                ccspvod_bl, 
                ccscva_bl, 
                ccsdemen_bl, 
                ccscopd_bl, 
                ccscontis_bl, 
                ccspud_bl, 
                # ccsmld_bl, # only have 0
                ccsdm_bl, 
                ccshemi_bl, 
                ccsmsrd_bl, 
                ccsdmeod_bl, 
                ccstumor_bl, 
                # ccsleuk_bl, # only have 0
                # ccslymph_bl, # only have 0
                ccsmsld_bl, 
                ccsmst_bl, 
                # ccsaids_bl # only have 0
                ) %>%
  tbl_summary(by = persistence, # stratify by outcome
              statistic = list(all_continuous() ~ "{mean} ({sd})", #use mean and standard deviation for continuous variables
                               all_categorical() ~ "{n} ({p}%)"), #use count and percentage for categorical variables
              digits = list(all_categorical() ~ c(0, 2), #adds no decimal places to counts and one decimal to percentages
                            all_continuous() ~ 1), #adds one decimal place to mean and sd values
              missing = "no"
              #missing_text = "Missing (n)"
              ) %>%
  add_p(pvalue_fun = ~style_pvalue(.x, digits = 2)) %>% #adds p value
  add_overall() %>% #adds column with non-stratified summary statistics
  modify_header(label ~ "**Variable**") %>% #the asterisks bold the label
  modify_spanning_header(c("stat_1", "stat_2") ~ "**Persistence**") %>%
  bold_labels() %>%
  as_gt() %>% 
  tab_header(
    title = "Table 4. Variables associated with MRSA persistence in new acquisition patient"
    ) %>% 
  tab_style(cell_borders(
    sides = c("top"), style = "hidden"),
    locations = cells_title())

# View the table
mrsa_uni_a
```
## VRE Univariate_Acquisition
```{r}
#| label: VRE univariate analysis_acquisition
# Colonization VRE Univariate Table (Crude)
vre_uni_a <- vre_unique_a %>%
  dplyr::select(age_bl,
                BMI,
                gender,
                race,
                vre_p_cat,
                vre_env_cat,
                # ethnic_bl,# 1 only have 1 observation
                antibiotic_bl,
                deviceuse, 
                feedtube,
                urincath, 
                openwound_final,
                cogsum_bl,
                hovre_bl,
                psmscore,
                charlson_cat,
                hospstay,
                visit_intervened,
                persistence,
                ccsmi_bl, 
                ccschf_bl, 
                ccspvod_bl, 
                ccscva_bl, 
                ccsdemen_bl, 
                ccscopd_bl, 
                ccscontis_bl, 
                # ccspud_bl, # only have 0
                # ccsmld_bl, # only have 0
                ccsdm_bl, 
                ccshemi_bl, 
                ccsmsrd_bl, 
                ccsdmeod_bl, 
                ccstumor_bl, 
                ccsleuk_bl, 
                # ccslymph_bl,# only have 0
                ccsmsld_bl, 
                ccsmst_bl, 
                # ccsaids_bl # only have 0
                ) %>%
  tbl_summary(by = persistence, # stratify by outcome
              statistic = list(all_continuous() ~ "{mean} ({sd})", #use mean and standard deviation for continuous variables
                               all_categorical() ~ "{n} ({p}%)"), #use count and percentage for categorical variables
              digits = list(all_categorical() ~ c(0, 2), #adds no decimal places to counts and one decimal to percentages
                            all_continuous() ~ 1), #adds one decimal place to mean and sd values
              missing = "no"
              #missing_text = "Missing (n)"
              ) %>%
  add_p(pvalue_fun = ~style_pvalue(.x, digits = 2)) %>% #adds p value
  add_overall() %>% #adds column with non-stratified summary statistics
  modify_header(label ~ "**Variable**") %>% #the asterisks bold the label
  modify_spanning_header(c("stat_1", "stat_2") ~ "**Persistence**") %>%
  bold_labels() %>%
  as_gt() %>% 
  tab_header(
    title = "Table 4. Variables associated with VRE persistence in new acquisition patient"
    ) %>% 
  tab_style(cell_borders(
    sides = c("top"), style = "hidden"),
    locations = cells_title())

# View the table
vre_uni_a
```

# Table 5: Multivariate for Persistence
## Multivariate - Colonization
```{r}
#| label: multivariate analysis_colonization
# Multivariate logistic models
model_rgnb_c <- glm(persistence ~ age_bl + gender + gnbr_p + openwound_final + cogsum_bl, 
                   data = rgnb_unique_c, 
                   family = binomial(link = "logit"))
summary(model_rgnb_c)

model_mrsa_c <- glm(persistence ~ age_bl + gender + mrsa_p, 
                   data = mrsa_unique_c, 
                   family = binomial(link = "logit"))
summary(model_mrsa_c)

model_vre_c <- glm(persistence ~ age_bl + gender + vre_p + ccspvod_bl, 
                   data = vre_unique_c, 
                   family = binomial(link = "logit"))
summary(model_vre_c)
```
## Multivariate - New Acquisition
```{r}
#| label: multivariate analysis_New acquisition
# Multivariate logistic models
model_rgnb_a <- glm(persistence ~ age_bl + gender + deviceuse + urincath + ccshemi_bl + ccsmsrd_bl, 
                   data = rgnb_unique_a, 
                   family = binomial(link = "logit"))
summary(model_rgnb_a)

model_mrsa_a <- glm(persistence ~ age_bl + gender + hospstay, 
                   data = mrsa_unique_a, 
                   family = binomial(link = "logit"))
summary(model_mrsa_a)

model_vre_a <- glm(persistence ~ age_bl + BMI + gender + vre_p + ccschf_bl, 
                   data = vre_unique_a, 
                   family = binomial(link = "logit"))
summary(model_vre_a)
```

# Table 6: Univariate Analysis for Duration
## RGNB univairate analysis_duration_colonization
```{r}
#| label: RGNB univariate analysis_colonization_duration
```

# Figure 3.Survival analysis
## New Acquisition
```{r}
# Load necessary libraries
library(survival)
library(ranger)
library(ggplot2)
library(dplyr)
library(ggfortify)
library(survminer)

# Create subsets only have "persistent"
persistence_a <- subset(combined_unique_a) %>%
                         filter(persistence == "Persistent")

# Create a survival object
surv_a <- Surv(time = persistence_a$duration, 
             event = persistence_a$cleared == "Yes")

# Fit the Kaplan-Meier model
km_curve_a <- survfit(surv_a ~ organism, data = persistence_a)
summary(km_curve_a)
survdiff(surv_a ~ organism, data = persistence_a)

# Plot the Kaplan-Meier curve
gg_surv_a <- ggsurvplot(km_curve_a, 
           data = persistence_a,
           pval = TRUE,
           pval.method = TRUE,
           pval.coord = c(0, 0.1), # position the p-value on the plot
           pval.method.coord = c(0, 0.15), # position the method p-value on the plot
           ggtheme = theme_minimal(),
           # risk.table = TRUE,       # To show the risk table
           # risk.table.col = "strata",# Change risk table color by groups
           legend.labs = c("MRSA", "RGNB", "VRE"), # Change group labels
           legend.title = "Organism",
           xlab = "Clerance of colonization (days)",
           ylab = "Probability of MDROs Persistnece",
           title = "Kaplan-Meier Survival Curve for New Acquisition Residents"
)

# Calculate median survival times
median_surv_a <- surv_median(km_curve_a)
median_surv_a

# Add median survival times to the plot
# medians <- median_surv$median
#for(i in 1:nrow(median_surv)) {
#  gg_surv$plot <- gg_surv$plot +
#    annotate("text", x = medians[i], y = 0.5, label = #paste("Median:", round(medians[i], 2)),
#             color = "black", vjust = -1) # Adjust color and vjust as needed
#}

# Extract survival probabilities at specific times (e.g., 50 and 100 days)
# surv_probs <- summary(km_curve_a, times = c(50, 100))
# print(surv_probs)

# Add survival probabilities to the plot
# for (i in 1:length(km_curve_a$strata)) {
#  organism_name <- names(km_curve_a$strata)[i]
#  prob_50 <- surv_probs$surv[which(surv_probs$strata == organism_name & surv_probs$time == 50)]
#  prob_100 <- surv_probs$surv[which(surv_probs$strata == organism_name & surv_probs$time == 100)]

#  gg_surv$plot <- gg_surv$plot +
 #   annotate("text", x = 50, y = prob_50, label = paste(organism_name, "50 days:", round(prob_50, 2)),
  #           color = "black", vjust = -1, hjust = 0) +
   # annotate("text", x = 100, y = prob_100, label = paste(organism_name, "100 days:", round(prob_100, 2)),
    #         color = "black", vjust = -1, hjust = 0)
#}

# Print the plot
print(gg_surv_a)

# ggsave(file = "survival_1.png", print(gg_surv_a), width = 8, height = 6, dpi = 600)
```
### Split Acquisition
```{r}
# Create subsets for each organism
persistence_rgnb_a <- subset(persistence_a, organism == "RGNB")
persistence_mrsa_a <- subset(persistence_a, organism == "MRSA")
persistence_vre_a <- subset(persistence_a, organism == "VRE")

# Kaplan-Meier analysis for RGNB
surv_rgnb_a <- Surv(time = persistence_rgnb_a$duration, event = persistence_rgnb_a$cleared == "Yes")
km_curve_rgnb_a <- survfit(surv_rgnb_a ~ 1)
summary(km_curve_rgnb_a)
ggsurvplot(km_curve_rgnb_a, 
           data = persistence_rgnb_a,
           ggtheme = theme_minimal(),
           xlab = "Clearance of colonization (days)",
           ylab = "Probability of MDROs Persistence",
           title = "Kaplan-Meier Survival Curve for RGNB Residents",
           legend = "none"
)

# Kaplan-Meier analysis for MRSA
surv_mrsa_a <- Surv(time = persistence_mrsa_a$duration, event = persistence_mrsa_a$cleared == "Yes")
km_curve_mrsa_a <- survfit(surv_mrsa_a ~ 1)
summary(km_curve_mrsa_a)
ggsurvplot(km_curve_mrsa_a, 
           data = persistence_mrsa_a,
           ggtheme = theme_minimal(),
           xlab = "Clearance of colonization (days)",
           ylab = "Probability of MDROs Persistence",
           title = "Kaplan-Meier Survival Curve for MRSA Residents",
           legend = "none"
)

# Kaplan-Meier analysis for VRE
surv_vre_a <- Surv(time = persistence_vre_a$duration, event = persistence_vre_a$cleared == "Yes")
km_curve_vre_a <- survfit(surv_vre_a ~ 1)
summary(km_curve_vre_a)
ggsurvplot(km_curve_vre_a, 
           data = persistence_vre_a,
           ggtheme = theme_minimal(),
           xlab = "Clearance of colonization (days)",
           ylab = "Probability of MDROs Persistence",
           title = "Kaplan-Meier Survival Curve for VRE Residents",
           legend = "none"
)

# Cox proportional hazards model for RGNB
cox_rgnb_a <- coxph(surv_rgnb_a ~ age_bl + gender, data = persistence_rgnb_a)
summary(cox_rgnb_a)
exp(cbind(coef(cox_rgnb_a), confint(cox_rgnb_a)))
cox.zph(cox_rgnb_a)
plot(cox.zph(cox_rgnb_a))

# Cox proportional hazards model for VRE
cox_vre_a <- coxph(surv_vre_a ~ age_bl + gender, data = persistence_vre_a)
summary(cox_vre_a)
exp(cbind(coef(cox_vre_a), confint(cox_vre_a)))
cox.zph(cox_vre_a)
plot(cox.zph(cox_vre_a))
```

## Colonization
```{r}
# Load necessary libraries
library(survival)
library(ranger)
library(ggplot2)
library(dplyr)
library(ggfortify)
library(survminer)

# Create subsets only have "persistent"
persistence_c <- subset(combined_unique_c) %>%
                         filter(persistence == "Persistent")

# Create a survival object
surv_c <- Surv(time = persistence_c$duration, 
             event = persistence_c$cleared == "Yes")

# Fit the Kaplan-Meier model
km_curve_c <- survfit(surv_c ~ organism, data = persistence_c)
summary(km_curve_c)
survdiff(surv_c ~ organism, data = persistence_c)

# Plot the Kaplan-Meier curve
gg_surv_c <- ggsurvplot(km_curve_c, 
           data = persistence_c,
           pval = TRUE,
           pval.method = TRUE,
           pval.coord = c(0, 0.1), # position the p-value on the plot
           pval.method.coord = c(0, 0.15), # position the method p-value on the plot
           ggtheme = theme_minimal(),
           # conf.int = TRUE,
           # risk.table = TRUE,       # To show the risk table
           # risk.table.col = "strata",# Change risk table color by groups
           legend.labs = c("MRSA", "RGNB", "VRE"), # Change group labels
           legend.title = "Organism",
           xlab = "Clerance of colonization (days)",
           ylab = "Probability of MDROs Persistnece",
           title = "Kaplan-Meier Survival Curve for Preexisting Colonization Residents"
)

# Calculate median survival times
median_surv_c <- surv_median(km_curve_c)
median_surv_c

# Add median survival times to the plot
# medians <- median_surv$median
#for(i in 1:nrow(median_surv)) {
#  gg_surv$plot <- gg_surv$plot +
#    annotate("text", x = medians[i], y = 0.5, label = #paste("Median:", round(medians[i], 2)),
#             color = "black", vjust = -1) # Adjust color and vjust as needed
#}

# Extract survival probabilities at specific times (e.g., 50 and 100 days)
# surv_probs <- summary(km_curve_a, times = c(50, 100))
# print(surv_probs)

# Add survival probabilities to the plot
# for (i in 1:length(km_curve_a$strata)) {
#  organism_name <- names(km_curve_a$strata)[i]
#  prob_50 <- surv_probs$surv[which(surv_probs$strata == organism_name & surv_probs$time == 50)]
#  prob_100 <- surv_probs$surv[which(surv_probs$strata == organism_name & surv_probs$time == 100)]

#  gg_surv$plot <- gg_surv$plot +
 #   annotate("text", x = 50, y = prob_50, label = paste(organism_name, "50 days:", round(prob_50, 2)),
  #           color = "black", vjust = -1, hjust = 0) +
   # annotate("text", x = 100, y = prob_100, label = paste(organism_name, "100 days:", round(prob_100, 2)),
    #         color = "black", vjust = -1, hjust = 0)
#}

# Print the plot
print(gg_surv_c)

# Cox proportional hazard
cox_c <- coxph(surv_c ~ organism, data = persistence_c)
summary(cox_c)
# Hazard Ratios (95% CI)
exp(cbind(coef(cox_c),coefci(cox_c)))
cox.zph(cox_c) # tests whether the hazards are proportional each variable separately
plot(cox.zph(cox_c)) # Check assumptions
```
### Split Colonization
```{r}
persistence_rgnb_c <- subset(persistence_c, organism == "RGNB")
persistence_mrsa_c <- subset(persistence_c, organism == "MRSA")
persistence_vre_c <- subset(persistence_c, organism == "VRE")

# Kaplan-Meier analysis for RGNB
surv_rgnb_c <- Surv(time = persistence_rgnb_c$duration, event = persistence_rgnb_c$cleared == "Yes")
km_curve_rgnb_c <- survfit(surv_rgnb_c ~ 1)
summary(km_curve_rgnb_c)
ggsurvplot(km_curve_rgnb_c, 
           data = persistence_rgnb_c,
           ggtheme = theme_minimal(),
           xlab = "Clearance of colonization (days)",
           ylab = "Probability of MDROs Persistence",
           title = "Kaplan-Meier Survival Curve for RGNB Residents",
           legend = "none"
)

# Kaplan-Meier analysis for MRSA
surv_mrsa_c <- Surv(time = persistence_mrsa_c$duration, event = persistence_mrsa_c$cleared == "Yes")
km_curve_mrsa_c <- survfit(surv_mrsa_c ~ 1)
summary(km_curve_mrsa_c)
ggsurvplot(km_curve_mrsa_c, 
           data = persistence_mrsa_c,
           ggtheme = theme_minimal(),
           xlab = "Clearance of colonization (days)",
           ylab = "Probability of MDROs Persistence",
           title = "Kaplan-Meier Survival Curve for MRSA Residents",
           legend = "none"
)

# Kaplan-Meier analysis for VRE
surv_vre_c <- Surv(time = persistence_vre_c$duration, event = persistence_vre_c$cleared == "Yes")
km_curve_vre_c <- survfit(surv_vre_c ~ 1)
summary(km_curve_vre_c)
ggsurvplot(km_curve_vre_c, 
           data = persistence_vre_c,
           ggtheme = theme_minimal(),
           xlab = "Clearance of colonization (days)",
           ylab = "Probability of MDROs Persistence",
           title = "Kaplan-Meier Survival Curve for VRE Residents",
           legend = "none"
)

# Cox proportional hazards model for RGNB
cox_rgnb_c <- coxph(surv_rgnb_c ~ age_bl + gender, data = persistence_rgnb_c)
summary(cox_rgnb_c)
exp(cbind(coef(cox_rgnb_c), confint(cox_rgnb_c)))
cox.zph(cox_rgnb_c)
plot(cox.zph(cox_rgnb_c))

# Cox proportional hazards model for MRSA
cox_mrsa_c <- coxph(surv_mrsa_c ~ age_bl + gender, data = persistence_mrsa_c)
summary(cox_mrsa_c)
exp(cbind(coef(cox_mrsa_c), confint(cox_mrsa_c)))
cox.zph(cox_mrsa_c)
plot(cox.zph(cox_mrsa_c))

# Cox proportional hazards model for VRE
cox_vre_c <- coxph(surv_vre_c ~ age_bl + gender, data = persistence_vre_c)
summary(cox_vre_c)
exp(cbind(coef(cox_vre_c), confint(cox_vre_c)))
cox.zph(cox_vre_c)
plot(cox.zph(cox_vre_c))
```



