---
title: "Pathways Persistence of MDROs Analysis"
author: "Skylar Zeng"
date: "Last compiled on `r format(Sys.Date(), '%B %d, %Y')`"
format:
  docx:
    code-overflow: wrap
embed-resources: true
---

# Install and Load Libraries
```{r}
#| label: install_packages
#| eval: false
install.packages("readr")
install.packages("readxl")
install.packages("dplyr")
install.packages("tidyr")
install.packages("gtsummary")
install.packages("gt")
install.packages("sjlabelled")
install.packages("writexl")
install.packages("parameters")
install.packages("bayestestR")
install.packages("purrr")
install.packages("ggplot2")
install.packages("ggpubr")
install.packages("survival")
install.packages("ranger")
install.packages("ggfortify")
install.packages("survminer")
install.packages("lmtest")
install.packages("labelled")
```

```{r}
#| label: load_packages
#| message: false
library(readr)
library(readxl)
library(dplyr)
library(tidyr)
library(gtsummary)
library(gt)
library(sjlabelled)
library(writexl)
library(parameters)
library(bayestestR)
library(purrr)
library(ggplot2)
library(ggpubr)
library(survival)
library(ranger)
library(ggfortify)
library(survminer)
library(lmtest)
library(labelled)
```

# Import Datasets
```{r}
#| label: import_dataset
#| message: false
Pathways_complete <- read_excel("pathwaysbothphases_0723.xlsx")
Pathways_dictionary <- read_excel("Pathways Database Dictionary_Derived Vars in second sheet.xlsx")
```
## Subsets with Necessary Variables
```{r}
Pathway <- Pathways_complete[, c(
    "facilID", 
    "studyID",
    "visit", 
    "novisits", 
    "phase", 
    "newinfect",
    "maxvisit",
    "age_bl", 
    "sex_bl",
    "height_bl",
    "weight_bl",
    "race_bl",
    "ethnic_bl",
    "dtvisit", 
    "intervention",
    "antibiotic_bl", 
    "feedtube", 
    "urincath", 
    "deviceuse", 
    "openwound_final",
    "cogsum_bl",
    "homrsa_bl",
    "hovre_bl",
    "horgnb_bl",
    "psmscore",
    "anyanypos_p",
    "anymrsa_p",
    "anyvre_p",
    "anygnbr_p",
    "acqmrsa", 
    "acqvre", 
    "acqgnbr", 
    "hospstay",
    "charlson", 
    "ccsmi_bl", 
    "ccschf_bl", 
    "ccspvod_bl", 
    "ccscva_bl", 
    "ccsdemen_bl", 
    "ccscopd_bl", 
    "ccscontis_bl", 
    "ccspud_bl", 
    "ccsmld_bl", 
    "ccsdm_bl", 
    "ccshemi_bl", 
    "ccsmsrd_bl", 
    "ccsdmeod_bl", 
    "ccstumor_bl", 
    "ccsleuk_bl", 
    "ccslymph_bl", 
    "ccsmsld_bl", 
    "ccsmst_bl",
    "ccsaids_bl",
    "VREsiteE1",
    "VREsiteE10",
    "VREsiteE11",
    "VREsiteE12",
    "VREsiteE2",
    "VREsiteE3", 
    "VREsiteE4",
    "VREsiteE5",
    "VREsiteE6",
    "VREsiteE7",
    "VREsiteE8",
    "VREsiteE9",
    "VREsiteF",
    "VREsiteG",
    "VREsiteH",
    "VREsiteN",
    "VREsiteO",
    "VREsiteR",
    "VREsiteS",
    "vre_p",
    "MRSAsiteE1",
    "MRSAsiteE10",
    "MRSAsiteE11",
    "MRSAsiteE12",
    "MRSAsiteE2",
    "MRSAsiteE3", 
    "MRSAsiteE4",
    "MRSAsiteE5",
    "MRSAsiteE6",
    "MRSAsiteE7",
    "MRSAsiteE8",
    "MRSAsiteE9",
    "MRSAsiteF",
    "MRSAsiteG",
    "MRSAsiteH",
    "MRSAsiteN",
    "MRSAsiteO",
    "MRSAsiteR",
    "MRSAsiteS",
    "mrsa_p",
    "RGNBsiteE1",
    "RGNBsiteE10",
    "RGNBsiteE11",
    "RGNBsiteE12",
    "RGNBsiteE2",
    "RGNBsiteE3",
    "RGNBsiteE4",
    "RGNBsiteE5",
    "RGNBsiteE6",
    "RGNBsiteE7",
    "RGNBsiteE8",
    "RGNBsiteE9",
    "RGNBsiteF",
    "RGNBsiteG",
    "RGNBsiteH",
    "RGNBsiteN",
    "RGNBsiteO",
    "RGNBsiteR",
    "RGNBsiteS",
    "gnbr_p",
    "gnbr_env",
    "mrsa_env",
    "vre_env",
    "mrsa_acq_totvisits_r",
    "mrsa_acq_persistence",
    "mrsa_acq_cleared",
    "mrsa_acq_duration",
    "mrsa_pre_totvisits_r",
    "mrsa_pre_persistence",
    "mrsa_pre_cleared",
    "mrsa_pre_duration",
    "vre_acq_totvisits_r",
    "vre_acq_persistence",
    "vre_acq_cleared",
    "vre_acq_duration",
    "vre_pre_totvisits_r",
    "vre_pre_persistence",
    "vre_pre_cleared",
    "vre_pre_duration",
    "rgnb_acq_totvisits_r",
    "rgnb_acq_persistence",
    "rgnb_acq_cleared",
    "rgnb_acq_duration",
    "rgnb_pre_totvisits_r",
    "rgnb_pre_persistence",
    "rgnb_pre_cleared",
    "rgnb_pre_duration"
)]
```

## Change labels and numbers (99 & 999)
```{r}
# Function to replace "99" and "999" with NA in a vector
replace_with_na <- function(x) {
  x <- as.character(x)  # Convert to character to handle "99" and "999"
  x[x == "99"] <- NA
  x[x == "999"] <- NA
  return(x)
}

# Apply the function to all columns of the dataframe
Pathway[] <- lapply(Pathway, replace_with_na)

# Create BMI variable
Pathway <- Pathway %>%
  mutate(BMI = (as.numeric(weight_bl) * 703) / (as.numeric(height_bl)^2))

# Create visit_intervened variable (Intervention arm are all pnts in facility 1,2,6; interventions occurred in Phase 2)
Pathway <- Pathway %>%
  mutate(visit_intervened = case_when(
    phase == 1 ~ 0,
    phase == 2 & intervention == 0 ~ 0,
    phase == 2 & intervention == 1 ~ 1,
  ))

# Create new variable - Sex
Pathway <- Pathway %>% 
  mutate(Sex = case_when(sex_bl == 1 ~ "Male",
                         sex_bl == 0 ~ "Female"))

# Create new variable - Race
Pathway <- Pathway %>% 
  mutate(race_bl = case_when(race_bl == 1 ~ 0,
                          race_bl == 2 ~ 1))
Pathway <- Pathway %>% 
  mutate(Race = case_when(race_bl == 0 ~ "White",
                          race_bl == 1 ~ "Black or African American"))

# Create new categorical variable for charlson score with <=3 and >3 classification
Pathway$charlson_cat <- ifelse(Pathway$charlson < 3, "<3", ">=3")

# Create new binary categorical variable for Body Sites colonized
Pathway$gnbr_p_cat <- ifelse(Pathway$gnbr_p < 2, "<2", ">=2")
Pathway$mrsa_p_cat <- ifelse(Pathway$mrsa_p < 2, "<2", ">=2")
Pathway$vre_p_cat <- ifelse(Pathway$vre_p < 2, "<2", ">=2")

# Create new binary categorical variable for hospital stay
Pathway$hospstay_cat <- ifelse(Pathway$hospstay < 14, "<14", ">=14")

# Replace persistence and cleared values with labels
library(dplyr)

# Create new categorical variable for Body Sites colonized
# Create new binary categorical variable for Body Sites colonized
Pathway$gnbr_env_cat <- ifelse(Pathway$gnbr_env == 0, "0", ">=1")
Pathway$mrsa_env_cat <- ifelse(Pathway$mrsa_env == 0, "0", ">=1")
Pathway$vre_env_cat <- ifelse(Pathway$vre_env == 0, "0", ">=1")

# Create new binary variable for Diabetes
Pathway$Diabetes <- ifelse(Pathway$ccsdmeod_bl == 1 | Pathway$ccsdm_bl == 1, 1, 0)
Pathway$Diabetes <- as.character(Pathway$Diabetes)

# Create a new variable for presence of comorbidities
Pathway <- Pathway %>%
  mutate(comorbidity_category = case_when(
    ccsmi_bl == 1 ~ "Myocardial Infarction",
    ccschf_bl == 1 ~ "Congestive Heart Failure",
    ccspvod_bl == 1 ~ "Peripheral Vascular Dx",
    ccscva_bl == 1 ~ "Cerebral Vascular Accident",
    ccsdemen_bl == 1 ~ "Dementia",
    ccscopd_bl == 1 ~ "COPD",
    ccscontis_bl == 1 ~ "Connective Tissue Dx",
    ccspud_bl == 1 ~ "Peptic Ulcer Dx",
    ccsmld_bl == 1 ~ "Mild Liver Dx",
    Diabetes == 1 ~ "Diabetes",
    ccshemi_bl == 1 ~ "Hemiplegia",
    ccsmsrd_bl == 1 ~ "Mod-Severe Renal Dx",
    ccstumor_bl == 1 ~ "Any Tumor",
    ccsleuk_bl == 1 ~ "Leukemia",
    ccslymph_bl == 1 ~ "Lymphoma",
    ccsmst_bl == 1 ~ "Metastatic Solid Tumor",
    ccsaids_bl == 1 ~ "AIDS",
    TRUE ~ "None",
  ))

# Only use Phase 1 for now
Pathway_1 <- Pathway %>%
  filter(phase == 1)

# Filter observations without intervention
Pathway_1_n_intervention <- Pathway_1 %>%
  filter(visit_intervened == 0)

# Calculate total residents number
Pathway_1_n_intervention_unique <- Pathway_1_n_intervention %>%
  distinct(studyID, .keep_all = TRUE)

# Calculate the median of psmscore
psmscore_median <- median(Pathway_1_n_intervention_unique$psmscore, na.rm = TRUE)
psmscore_median
# Create the new binary categorical variable
Pathway$psmscore_cat <- ifelse(Pathway$psmscore < psmscore_median, "<14", ">=14")

# Filter observations without intervention
Pathway_1_n_intervention <- Pathway_1 %>%
  filter(visit_intervened == 0)

# Calculate total residents number
Pathway_1_n_intervention_unique <- Pathway_1_n_intervention %>%
  distinct(studyID, .keep_all = TRUE)

# Set labels for the specified variables
Pathway <- Pathway %>%
  set_variable_labels(
    age_bl = "Age",
    Sex = "Sex",
    Race = "Race",
    antibiotic_bl = "Antibiotic Use",
    charlson = "Charlson comorbidity index score",
    charlson_cat = "Charlson comorbidity index score",
    feedtube = "Feeding Tube Use",
    urincath = "Urinary Catheter Use",
    deviceuse = "Device use",
    cogsum_bl = "Cognitive Sum Score",
    homrsa_bl = "History of MRSA",
    hovre_bl = "History of VRE",
    horgnb_bl = "History of RGNB",
    psmscore_cat = "Functional disability (PSMS Total Score)",
    hospstay_cat = "Length of Hospitalization",
    ccsmi_bl = "Myocardial Infarction",
    ccschf_bl = "Congestive Heart Failure",
    ccspvod_bl = "Peripheral Vascular Dx",
    ccscva_bl = "Cerebral Vascular Accident",
    ccsdemen_bl = "Dementia",
    ccscopd_bl = "COPD",
    ccscontis_bl = "Connective Tissue Dx",
    ccspud_bl = "Peptic Ulcer Dx",
    ccsmld_bl = "Mild Liver Dx",
    Diabetes = "Diabetes",
    ccshemi_bl = "Hemiplegia",
    ccsmsrd_bl = "Mod-Severe Renal Dx",
    ccstumor_bl = "Any Tumor",
    ccsleuk_bl = "Leukemia",
    ccslymph_bl = "Lymphoma",
    ccsmsld_bl = "Mod-Severe Liver Dx",
    ccsmst_bl = "Metastatic Solid Tumor",
    ccsaids_bl = "AIDS",
    openwound_final = "Wound",
    visit_intervened = "Intervention",
    comorbidity_category = "Presence of Comorbidities",
    gnbr_env_cat = "Number of Room Surfaces Colonized",
    mrsa_env_cat = "Number of Room Surfaces Colonized",
    vre_env_cat = "Number of Room Surfaces Colonized",
    gnbr_p_cat = "Number of patient sites colonization",
    mrsa_p_cat = "Number of patient sites colonization",
    vre_p_cat = "Number of patient sites colonization",
    rgnb_pre_persistence = "Persistence",
    mrsa_pre_persistence = "Persistence",
    vre_pre_persistence = "Persistence",
    rgnb_acq_persistence = "Persistence",
    mrsa_acq_persistence = "Persistence",
    vre_acq_persistence = "Persistence",
    rgnb_pre_cleared = "Clearance",
    mrsa_pre_cleared = "Clearance",
    vre_pre_cleared = "Clearance",
    rgnb_acq_cleared = "Clearance",
    mrsa_acq_cleared = "Clearance",
    vre_acq_cleared = "Clearance",
  )
```

# Subset to different organisms
```{r}
# Subset datasets based on the presence of numbers in the persistence columns
rgnb_pre <- subset(Pathway_1_n_intervention, !is.na(as.numeric(as.character(rgnb_pre_persistence))) & rgnb_pre_persistence != "")
mrsa_pre <- subset(Pathway_1_n_intervention, !is.na(as.numeric(as.character(mrsa_pre_persistence))) & mrsa_pre_persistence != "")
vre_pre <- subset(Pathway_1_n_intervention, !is.na(as.numeric(as.character(vre_pre_persistence))) & vre_pre_persistence != "")

rgnb_acq <- subset(Pathway_1_n_intervention, !is.na(as.numeric(as.character(rgnb_acq_persistence))) & rgnb_acq_persistence != "")
mrsa_acq <- subset(Pathway_1_n_intervention, !is.na(as.numeric(as.character(mrsa_acq_persistence))) & mrsa_acq_persistence != "")
vre_acq <- subset(Pathway_1_n_intervention, !is.na(as.numeric(as.character(vre_acq_persistence))) & vre_acq_persistence != "")
```

## Colonization Subsets with different MDROs
```{r}
rgnb_pre <- rgnb_pre[, c(
    "facilID", 
    "studyID",
    "visit", 
    "novisits", 
    "phase", 
    "maxvisit",
    "age_bl", 
    "Sex",
    "height_bl",
    "weight_bl",
    "Race",
    "ethnic_bl",
    "dtvisit",
    "antibiotic_bl", 
    "feedtube", 
    "urincath", 
    "deviceuse", 
    "openwound_final",
    "cogsum_bl",
    "horgnb_bl",
    "psmscore",
    "anyanypos_p",
    "anygnbr_p",
    "acqgnbr", 
    "hospstay", 
    "charlson", 
    "ccsmi_bl", 
    "ccschf_bl", 
    "ccspvod_bl", 
    "ccscva_bl", 
    "ccsdemen_bl", 
    "ccscopd_bl", 
    "ccscontis_bl", 
    "ccspud_bl", 
    "ccsmld_bl", 
    "ccsdm_bl", 
    "ccshemi_bl", 
    "ccsmsrd_bl", 
    "ccsdmeod_bl", 
    "ccstumor_bl", 
    "ccsleuk_bl", 
    "ccslymph_bl", 
    "ccsmsld_bl",
    "ccsmst_bl", 
    "ccsaids_bl",
    "comorbidity_category",
    "RGNBsiteE1",
    "RGNBsiteE10",
    "RGNBsiteE11",
    "RGNBsiteE12",
    "RGNBsiteE2",
    "RGNBsiteE3",
    "RGNBsiteE4",
    "RGNBsiteE5",
    "RGNBsiteE6",
    "RGNBsiteE7",
    "RGNBsiteE8",
    "RGNBsiteE9",
    "RGNBsiteF",
    "RGNBsiteG",
    "RGNBsiteH",
    "RGNBsiteN",
    "RGNBsiteO",
    "RGNBsiteR",
    "RGNBsiteS",
    "gnbr_p",
    "gnbr_p_cat",
    "BMI",
    "visit_intervened",
    "charlson_cat",
    "gnbr_env",
    "gnbr_env_cat",
    "rgnb_pre_totvisits_r",
    "rgnb_pre_persistence",
    "rgnb_pre_cleared",
    "rgnb_pre_duration",
    "psmscore_cat",
    "hospstay_cat",
    "Diabetes"
)]

mrsa_pre <- mrsa_pre[, c(
    "facilID", 
    "studyID",
    "visit", 
    "novisits", 
    "phase", 
    "maxvisit",
    "age_bl", 
    "Sex",
    "height_bl",
    "weight_bl",
    "Race",
    "ethnic_bl",
    "dtvisit",
    "antibiotic_bl", 
    "feedtube", 
    "urincath", 
    "deviceuse", 
    "openwound_final",
    "cogsum_bl",
    "homrsa_bl",
    "psmscore",
    "anyanypos_p",
    "anymrsa_p",
    "acqmrsa", 
    "hospstay", 
    "charlson", 
    "ccsmi_bl", 
    "ccschf_bl", 
    "ccspvod_bl", 
    "ccscva_bl", 
    "ccsdemen_bl", 
    "ccscopd_bl", 
    "ccscontis_bl", 
    "ccspud_bl", 
    "ccsmld_bl", 
    "ccsdm_bl", 
    "ccshemi_bl", 
    "ccsmsrd_bl", 
    "ccsdmeod_bl", 
    "ccstumor_bl", 
    "ccsleuk_bl", 
    "ccslymph_bl", 
    "ccsmsld_bl",
    "ccsmst_bl", 
    "ccsaids_bl",
    "comorbidity_category",
    "MRSAsiteE1",
    "MRSAsiteE10",
    "MRSAsiteE11",
    "MRSAsiteE12",
    "MRSAsiteE2",
    "MRSAsiteE3",
    "MRSAsiteE4",
    "MRSAsiteE5",
    "MRSAsiteE6",
    "MRSAsiteE7",
    "MRSAsiteE8",
    "MRSAsiteE9",
    "MRSAsiteF",
    "MRSAsiteG",
    "MRSAsiteH",
    "MRSAsiteN",
    "MRSAsiteO",
    "MRSAsiteR",
    "MRSAsiteS",
    "mrsa_p",
    "mrsa_p_cat",
    "BMI",
    "visit_intervened",
    "charlson_cat",
    "mrsa_env",
    "mrsa_env_cat",
    "mrsa_pre_totvisits_r",
    "mrsa_pre_persistence",
    "mrsa_pre_cleared",
    "mrsa_pre_duration",
    "psmscore_cat",
    "hospstay_cat",
    "Diabetes"
)]

vre_pre <- vre_pre[, c(
    "facilID", 
    "studyID",
    "visit", 
    "novisits", 
    "phase", 
    "maxvisit",
    "age_bl", 
    "Sex",
    "height_bl",
    "weight_bl",
    "Race",
    "ethnic_bl",
    "dtvisit",
    "antibiotic_bl", 
    "feedtube", 
    "urincath", 
    "deviceuse", 
    "openwound_final",
    "cogsum_bl",
    "hovre_bl",
    "psmscore",
    "anyanypos_p",
    "anyvre_p",
    "acqvre", 
    "hospstay", 
    "charlson", 
    "ccsmi_bl", 
    "ccschf_bl", 
    "ccspvod_bl", 
    "ccscva_bl", 
    "ccsdemen_bl", 
    "ccscopd_bl", 
    "ccscontis_bl", 
    "ccspud_bl", 
    "ccsmld_bl", 
    "ccsdm_bl", 
    "ccshemi_bl", 
    "ccsmsrd_bl", 
    "ccsdmeod_bl", 
    "ccstumor_bl", 
    "ccsleuk_bl", 
    "ccslymph_bl", 
    "ccsmsld_bl",
    "ccsmst_bl", 
    "ccsaids_bl",
    "comorbidity_category",
    "VREsiteE1",
    "VREsiteE10",
    "VREsiteE11",
    "VREsiteE12",
    "VREsiteE2",
    "VREsiteE3",
    "VREsiteE4",
    "VREsiteE5",
    "VREsiteE6",
    "VREsiteE7",
    "VREsiteE8",
    "VREsiteE9",
    "VREsiteF",
    "VREsiteG",
    "VREsiteH",
    "VREsiteN",
    "VREsiteO",
    "VREsiteR",
    "VREsiteS",
    "vre_p",
    "vre_p_cat",
    "BMI",
    "visit_intervened",
    "charlson_cat",
    "vre_env",
    "vre_env_cat",
    "vre_pre_totvisits_r",
    "vre_pre_persistence",
    "vre_pre_cleared",
    "vre_pre_duration",
    "psmscore_cat",
    "hospstay_cat",
    "Diabetes"
)]
```

## New Acquisition Subsets with different MDROs
```{r}
rgnb_acq <- rgnb_acq[, c(
    "facilID", 
    "studyID",
    "visit", 
    "novisits", 
    "phase", 
    "maxvisit",
    "age_bl", 
    "Sex",
    "height_bl",
    "weight_bl",
    "Race",
    "ethnic_bl",
    "dtvisit",
    "antibiotic_bl", 
    "feedtube", 
    "urincath", 
    "deviceuse", 
    "openwound_final",
    "cogsum_bl",
    "horgnb_bl",
    "psmscore",
    "anyanypos_p",
    "anygnbr_p",
    "acqgnbr", 
    "hospstay", 
    "charlson", 
    "ccsmi_bl", 
    "ccschf_bl", 
    "ccspvod_bl", 
    "ccscva_bl", 
    "ccsdemen_bl", 
    "ccscopd_bl", 
    "ccscontis_bl", 
    "ccspud_bl", 
    "ccsmld_bl", 
    "ccsdm_bl", 
    "ccshemi_bl", 
    "ccsmsrd_bl", 
    "ccsdmeod_bl", 
    "ccstumor_bl", 
    "ccsleuk_bl", 
    "ccslymph_bl", 
    "ccsmsld_bl",
    "ccsmst_bl", 
    "ccsaids_bl",
    "comorbidity_category",
    "RGNBsiteE1",
    "RGNBsiteE10",
    "RGNBsiteE11",
    "RGNBsiteE12",
    "RGNBsiteE2",
    "RGNBsiteE3",
    "RGNBsiteE4",
    "RGNBsiteE5",
    "RGNBsiteE6",
    "RGNBsiteE7",
    "RGNBsiteE8",
    "RGNBsiteE9",
    "RGNBsiteF",
    "RGNBsiteG",
    "RGNBsiteH",
    "RGNBsiteN",
    "RGNBsiteO",
    "RGNBsiteR",
    "RGNBsiteS",
    "gnbr_p",
    "gnbr_p_cat",
    "BMI",
    "visit_intervened",
    "charlson_cat",
    "gnbr_env",
    "gnbr_env_cat",
    "rgnb_acq_totvisits_r",
    "rgnb_acq_persistence",
    "rgnb_acq_cleared",
    "rgnb_acq_duration",
    "psmscore_cat",
    "hospstay_cat",
    "Diabetes"
)]

mrsa_acq <- mrsa_acq[, c(
    "facilID", 
    "studyID",
    "visit", 
    "novisits", 
    "phase", 
    "maxvisit",
    "age_bl", 
    "Sex",
    "height_bl",
    "weight_bl",
    "Race",
    "ethnic_bl",
    "dtvisit",
    "antibiotic_bl", 
    "feedtube", 
    "urincath", 
    "deviceuse", 
    "openwound_final",
    "cogsum_bl",
    "homrsa_bl",
    "psmscore",
    "anyanypos_p",
    "anymrsa_p",
    "acqmrsa", 
    "hospstay", 
    "charlson", 
    "ccsmi_bl", 
    "ccschf_bl", 
    "ccspvod_bl", 
    "ccscva_bl", 
    "ccsdemen_bl", 
    "ccscopd_bl", 
    "ccscontis_bl", 
    "ccspud_bl", 
    "ccsmld_bl", 
    "ccsdm_bl", 
    "ccshemi_bl", 
    "ccsmsrd_bl", 
    "ccsdmeod_bl", 
    "ccstumor_bl", 
    "ccsleuk_bl", 
    "ccslymph_bl", 
    "ccsmsld_bl",
    "ccsmst_bl", 
    "ccsaids_bl",
    "comorbidity_category",
    "MRSAsiteE1",
    "MRSAsiteE10",
    "MRSAsiteE11",
    "MRSAsiteE12",
    "MRSAsiteE2",
    "MRSAsiteE3",
    "MRSAsiteE4",
    "MRSAsiteE5",
    "MRSAsiteE6",
    "MRSAsiteE7",
    "MRSAsiteE8",
    "MRSAsiteE9",
    "MRSAsiteF",
    "MRSAsiteG",
    "MRSAsiteH",
    "MRSAsiteN",
    "MRSAsiteO",
    "MRSAsiteR",
    "MRSAsiteS",
    "mrsa_p",
    "mrsa_p_cat",
    "BMI",
    "visit_intervened",
    "charlson_cat",
    "mrsa_env",
    "mrsa_env_cat",
    "mrsa_acq_totvisits_r",
    "mrsa_acq_persistence",
    "mrsa_acq_cleared",
    "mrsa_acq_duration",
    "psmscore_cat",
    "hospstay_cat",
    "Diabetes"
)]

vre_acq <- vre_acq[, c(
    "facilID", 
    "studyID",
    "visit", 
    "novisits", 
    "phase", 
    "maxvisit",
    "age_bl", 
    "Sex",
    "height_bl",
    "weight_bl",
    "Race",
    "ethnic_bl",
    "dtvisit",
    "antibiotic_bl", 
    "feedtube", 
    "urincath", 
    "deviceuse", 
    "openwound_final",
    "cogsum_bl",
    "hovre_bl",
    "psmscore",
    "anyanypos_p",
    "anyvre_p",
    "acqvre", 
    "hospstay", 
    "charlson", 
    "ccsmi_bl", 
    "ccschf_bl", 
    "ccspvod_bl", 
    "ccscva_bl", 
    "ccsdemen_bl", 
    "ccscopd_bl", 
    "ccscontis_bl", 
    "ccspud_bl", 
    "ccsmld_bl", 
    "ccsdm_bl", 
    "ccshemi_bl", 
    "ccsmsrd_bl", 
    "ccsdmeod_bl", 
    "ccstumor_bl", 
    "ccsleuk_bl", 
    "ccslymph_bl", 
    "ccsmsld_bl",
    "ccsmst_bl", 
    "ccsaids_bl",
    "comorbidity_category",
    "VREsiteE1",
    "VREsiteE10",
    "VREsiteE11",
    "VREsiteE12",
    "VREsiteE2",
    "VREsiteE3",
    "VREsiteE4",
    "VREsiteE5",
    "VREsiteE6",
    "VREsiteE7",
    "VREsiteE8",
    "VREsiteE9",
    "VREsiteF",
    "VREsiteG",
    "VREsiteH",
    "VREsiteN",
    "VREsiteO",
    "VREsiteR",
    "VREsiteS",
    "vre_p",
    "vre_p_cat",
    "BMI",
    "visit_intervened",
    "charlson_cat",
    "vre_env",
    "vre_env_cat",
    "vre_acq_totvisits_r",
    "vre_acq_persistence",
    "vre_acq_cleared",
    "vre_acq_duration",
    "psmscore_cat",
    "hospstay_cat",
    "Diabetes"
)]
```

## Adjust subset and summary the number of subsets for different MDRO
```{r}
# Copy labels for rgnb_pre
rgnb_pre <- copy_labels(Pathway, rgnb_pre)
# Copy labels for rgnb_acq
rgnb_acq <- copy_labels(Pathway, rgnb_acq)
# Copy labels for mrsa_pre
mrsa_pre <- copy_labels(Pathway, mrsa_pre)
# Copy labels for mrsa_acq
mrsa_acq <- copy_labels(Pathway, mrsa_acq)
# Copy labels for vre_pre
vre_pre <- copy_labels(Pathway, vre_pre)
# Copy labels for vre_acq
vre_acq <- copy_labels(Pathway, vre_acq)

# Apply mutations to rgnb_pre
rgnb_pre <- rgnb_pre %>%
  mutate(rgnb_pre_persistence = factor(rgnb_pre_persistence, levels = c(0, 1), labels = c("Intermittent", "Persistent"))) %>%
  mutate(rgnb_pre_cleared = factor(rgnb_pre_cleared, levels = c(0, 1), labels = c("No", "Yes")))

# Apply mutations to rgnb_acq
rgnb_acq <- rgnb_acq %>%
  mutate(rgnb_acq_persistence = factor(rgnb_acq_persistence, levels = c(0, 1), labels = c("Intermittent", "Persistent"))) %>%
  mutate(rgnb_acq_cleared = factor(rgnb_acq_cleared, levels = c(0, 1), labels = c("No", "Yes")))

# Apply mutations to mrsa_pre
mrsa_pre <- mrsa_pre %>%
  mutate(mrsa_pre_persistence = factor(mrsa_pre_persistence, levels = c(0, 1), labels = c("Intermittent", "Persistent"))) %>%
  mutate(mrsa_pre_cleared = factor(mrsa_pre_cleared, levels = c(0, 1), labels = c("No", "Yes")))

# Apply mutations to mrsa_acq
mrsa_acq <- mrsa_acq %>%
  mutate(mrsa_acq_persistence = factor(mrsa_acq_persistence, levels = c(0, 1), labels = c("Intermittent", "Persistent"))) %>%
  mutate(mrsa_acq_cleared = factor(mrsa_acq_cleared, levels = c(0, 1), labels = c("No", "Yes")))

# Apply mutations to vre_pre
vre_pre <- vre_pre %>%
  mutate(vre_pre_persistence = factor(vre_pre_persistence, levels = c(0, 1), labels = c("Intermittent", "Persistent"))) %>%
  mutate(vre_pre_cleared = factor(vre_pre_cleared, levels = c(0, 1), labels = c("No", "Yes")))

# Apply mutations to vre_acq
vre_acq <- vre_acq %>%
  mutate(vre_acq_persistence = factor(vre_acq_persistence, levels = c(0, 1), labels = c("Intermittent", "Persistent"))) %>%
  mutate(vre_acq_cleared = factor(vre_acq_cleared, levels = c(0, 1), labels = c("No", "Yes")))

```

## Make unique datasets (studyID only appeared once)
```{r}
# Remove duplicate study IDs, keeping the one where acqgnbr = 1
rgnb_acq_unique <- rgnb_acq %>%
  group_by(studyID) %>%
  arrange(desc(acqgnbr)) %>%  # Sort in descending order, so 1 comes before 0
  slice(1) %>%  # Keep only the first row (which will be acqgnrb = 1 if it exists)
  ungroup()
# Remove duplicate study IDs, keeping the one where acqmrsa = 1
mrsa_acq_unique <- mrsa_acq %>%
  group_by(studyID) %>%
  arrange(desc(acqmrsa)) %>%  # Sort in descending order, so 1 comes before 0
  slice(1) %>%  # Keep only the first row (which will be acqgnrb = 1 if it exists)
  ungroup()
# Remove duplicate study IDs, keeping the one where acqgnbr = 1
vre_acq_unique <- vre_acq %>%
  group_by(studyID) %>%
  arrange(desc(acqvre)) %>%  # Sort in descending order, so 1 comes before 0
  slice(1) %>%  # Keep only the first row (which will be acqgnrb = 1 if it exists)
  ungroup()

# Combine Unique Table (New Acquisition)
combined_unique_a <- bind_rows(rgnb_acq_unique, mrsa_acq_unique, vre_acq_unique, .id = "organism")
combined_unique_a <- combined_unique_a %>%
  mutate(organism = case_when(
    organism == 1 ~ "RGNB",
    organism == 2 ~ "MRSA",
    organism == 3 ~ "VRE"
  ))

# Subset rgnb_pre where visit == 0 to make it unique
rgnb_pre_unique <- subset(rgnb_pre, visit == 0)
# Subset mrsa_pre where visit == 0 to make it unique
mrsa_pre_unique <- subset(mrsa_pre, visit == 0)
# Subset vre_pre where visit == 0 to make it unique
vre_pre_unique <- subset(vre_pre, visit == 0)

# Combine Unique Table (Preexisting Colonization)
combined_unique_c <- bind_rows(rgnb_pre_unique, mrsa_pre_unique, vre_pre_unique, .id = "organism")
combined_unique_c <- combined_unique_c %>%
  mutate(organism = case_when(
    organism == 1 ~ "RGNB",
    organism == 2 ~ "MRSA",
    organism == 3 ~ "VRE"
  ))
```
## Make Hands-only variable
```{r}
library(dplyr)

# Function to create "Hands only" variable based on conditions
create_hands_only <- function(df, siteH, siteF, siteG, siteN, siteO, siteR, siteS) {
  df %>%
    mutate(Hands_only = case_when(
      get(siteH) == 1 & (is.na(get(siteF)) | get(siteF) == 0) &
        (is.na(get(siteG)) | get(siteG) == 0) &
        (is.na(get(siteN)) | get(siteN) == 0) &
        (is.na(get(siteO)) | get(siteO) == 0) &
        (is.na(get(siteR)) | get(siteR) == 0) &
        (is.na(get(siteS)) | get(siteS) == 0) ~ 1,
      TRUE ~ 0
    ))
}

# Apply the function to each dataset
rgnb_pre_unique <- create_hands_only(rgnb_pre_unique, "RGNBsiteH", "RGNBsiteF", "RGNBsiteG", "RGNBsiteN", "RGNBsiteO", "RGNBsiteR", "RGNBsiteS")
rgnb_acq_unique <- create_hands_only(rgnb_acq_unique, "RGNBsiteH", "RGNBsiteF", "RGNBsiteG", "RGNBsiteN", "RGNBsiteO", "RGNBsiteR", "RGNBsiteS")

mrsa_pre_unique <- create_hands_only(mrsa_pre_unique, "MRSAsiteH", "MRSAsiteF", "MRSAsiteG", "MRSAsiteN", "MRSAsiteO", "MRSAsiteR", "MRSAsiteS")
mrsa_acq_unique <- create_hands_only(mrsa_acq_unique, "MRSAsiteH", "MRSAsiteF", "MRSAsiteG", "MRSAsiteN", "MRSAsiteO", "MRSAsiteR", "MRSAsiteS")

vre_pre_unique <- create_hands_only(vre_pre_unique, "VREsiteH", "VREsiteF", "VREsiteG", "VREsiteN", "VREsiteO", "VREsiteR", "VREsiteS")
vre_acq_unique <- create_hands_only(vre_acq_unique, "VREsiteH", "VREsiteF", "VREsiteG", "VREsiteN", "VREsiteO", "VREsiteR", "VREsiteS")
```

## Export datasets
```{r}
# Define the list of datasets and their corresponding filenames
datasets <- list(
  rgnb_pre = rgnb_pre,
  rgnb_acq = rgnb_acq,
  mrsa_pre = mrsa_pre,
  mrsa_acq = mrsa_acq,
  vre_pre = vre_pre,
  vre_acq = vre_acq,
  rgnb_pre_unique = rgnb_pre_unique,
  rgnb_acq_unique = rgnb_acq_unique,
  mrsa_pre_unique = mrsa_pre_unique,
  mrsa_acq_unique = mrsa_acq_unique,
  vre_pre_unique = vre_pre_unique,
  vre_acq_unique = vre_acq_unique
)

filenames <- c(
  "rgnb_pre_n_intervention.csv",
  "rgnb_acq_n_intervention.csv",
  "mrsa_pre_n_intervention.csv",
  "mrsa_acq_n_intervention.csv",
  "vre_pre_n_intervention.csv",
  "vre_acq_n_intervention.csv",
  "rgnb_pre_unique_n_intervention.csv",
  "rgnb_acq_unique_n_intervention.csv",
  "mrsa_pre_unique_n_intervention.csv",
  "mrsa_acq_unique_n_intervention.csv",
  "vre_pre_unique_n_intervention.csv",
  "vre_acq_unique_n_intervention.csv"
)

# Loop through the datasets and export each to a CSV file
for (i in seq_along(datasets)) {
  write.csv(datasets[[i]], file = filenames[i], row.names = FALSE)
}
```

# Table 1
## Rename columns for persistence and clearance in all datasets
```{r}
# Load the dplyr package
library(dplyr)

# Rename columns in the preexisting colonization datasets
rgnb_pre_unique <- rgnb_pre_unique %>%
  rename(Persistence = rgnb_pre_persistence, 
         Clearance = rgnb_pre_cleared,
         Duration = rgnb_pre_duration)

mrsa_pre_unique <- mrsa_pre_unique %>%
  rename(Persistence = mrsa_pre_persistence, 
         Clearance = mrsa_pre_cleared,
         Duration = mrsa_pre_duration)

vre_pre_unique <- vre_pre_unique %>%
  rename(Persistence = vre_pre_persistence, 
         Clearance = vre_pre_cleared,
         Duration = vre_pre_duration)

# Rename columns in the new acquisition datasets
rgnb_acq_unique <- rgnb_acq_unique %>%
  rename(Persistence = rgnb_acq_persistence, 
         Clearance = rgnb_acq_cleared,
         Duration = rgnb_acq_duration)

mrsa_acq_unique <- mrsa_acq_unique %>%
  rename(Persistence = mrsa_acq_persistence, 
         Clearance = mrsa_acq_cleared,
         Duration = mrsa_acq_duration)

vre_acq_unique <- vre_acq_unique %>%
  rename(Persistence = vre_acq_persistence, 
         Clearance = vre_acq_cleared,
         Duration = vre_acq_duration)
```
## Descriptive Statistics for each MDROs (both)
```{r}
# Summary for persistence in rgnb_pre_unique
summary_persistence_rgnb_pre <- rgnb_pre_unique %>%
  dplyr::select(Persistence, Clearance) %>%
  tbl_summary(
    type = list(Persistence ~ "categorical",
                Clearance ~ "categorical"),
    statistic = all_categorical() ~ "{n} ({p}%)",
    digits = list(all_categorical() ~ c(0, 1)),
    missing = "no"
  ) %>%
  modify_header(label ~ "**Variable**") %>%
  bold_labels() %>%
  modify_spanning_header(all_stat_cols() ~ "**RGNB**")

# Summary for persistence in mrsa_pre_unique
summary_persistence_mrsa_pre <- mrsa_pre_unique %>%
  dplyr::select(Persistence, Clearance) %>%
  tbl_summary(
    type = list(Persistence ~ "categorical",
                Clearance ~ "categorical"),
    statistic = all_categorical() ~ "{n} ({p}%)",
    digits = list(all_categorical() ~ c(0, 1)),
    missing = "no"
  ) %>%
  modify_header(label ~ "**Variable**") %>%
  bold_labels() %>%
  modify_spanning_header(all_stat_cols() ~ "**MRSA**")

# Summary for persistence in vre_pre_unique
summary_persistence_vre_pre <- vre_pre_unique %>%
  dplyr::select(Persistence, Clearance) %>%
  tbl_summary(
    type = list(Persistence ~ "categorical",
                Clearance ~ "categorical"),
    statistic = all_categorical() ~ "{n} ({p}%)",
    digits = list(all_categorical() ~ c(0, 1)),
    missing = "no"
  ) %>%
  modify_header(label ~ "**Variable**") %>%
  bold_labels() %>%
  modify_spanning_header(all_stat_cols() ~ "**VRE**")

# Combine summaries for persistence
combined_persistence_summary_pre <- tbl_merge(
  tbls = list(summary_persistence_rgnb_pre, summary_persistence_mrsa_pre, summary_persistence_vre_pre),
  tab_spanner = c("**RGNB**", "**MRSA**", "**VRE**")
) %>%
  as_gt() %>%
  tab_header(
    title = "Persistence and Clearance among NH Residents with Preexisting Colonization"
  ) %>%
  tab_style(
    cell_borders(sides = c("top"), style = "hidden"),
    locations = cells_title()
  )

# View and export the combined persistence table
combined_persistence_summary_pre

# Summary for persistence in rgnb_acq_unique
summary_persistence_rgnb_acq <- rgnb_acq_unique %>%
  dplyr::select(Persistence, Clearance) %>%
  tbl_summary(
    type = list(Persistence ~ "categorical",
                Clearance ~ "categorical"),
    statistic = all_categorical() ~ "{n} ({p}%)",
    digits = list(all_categorical() ~ c(0, 1)),
    missing = "no"
  ) %>%
  modify_header(label ~ "**Variable**") %>%
  bold_labels() %>%
  modify_spanning_header(all_stat_cols() ~ "**RGNB**")

# Summary for persistence in mrsa_acq_unique
summary_persistence_mrsa_acq <- mrsa_acq_unique %>%
  dplyr::select(Persistence, Clearance) %>%
  tbl_summary(
    type = list(Persistence ~ "categorical",
                Clearance ~ "categorical"),
    statistic = all_categorical() ~ "{n} ({p}%)",
    digits = list(all_categorical() ~ c(0, 1)),
    missing = "no"
  ) %>%
  modify_header(label ~ "**Variable**") %>%
  bold_labels() %>%
  modify_spanning_header(all_stat_cols() ~ "**MRSA**")

# Summary for persistence in vre_acq_unique
summary_persistence_vre_acq <- vre_acq_unique %>%
  dplyr::select(Persistence, Clearance) %>%
  tbl_summary(
    type = list(Persistence ~ "categorical",
                Clearance ~ "categorical"),
    statistic = all_categorical() ~ "{n} ({p}%)",
    digits = list(all_categorical() ~ c(0, 1)),
    missing = "no"
  ) %>%
  modify_header(label ~ "**Variable**") %>%
  bold_labels() %>%
  modify_spanning_header(all_stat_cols() ~ "**VRE**")

# Combine summaries for persistence
combined_persistence_summary_acq <- tbl_merge(
  tbls = list(summary_persistence_rgnb_acq, summary_persistence_mrsa_acq, summary_persistence_vre_acq),
  tab_spanner = c("**RGNB**", "**MRSA**", "**VRE**")
) %>%
  as_gt() %>%
  tab_header(
    title = "Persistence and Clearance among NH Residents with Acquired Colonization"
  ) %>%
  tab_style(
    cell_borders(sides = c("top"), style = "hidden"),
    locations = cells_title()
  )

# View and export the combined persistence table
combined_persistence_summary_acq
```
## Revised version (clearance >= 5)
```{r}
library(dplyr)
library(gtsummary)

# Filter the datasets
rgnb_pre_unique_filtered <- rgnb_pre_unique %>%
  filter(rgnb_pre_totvisits_r > 5)

mrsa_pre_unique_filtered <- mrsa_pre_unique %>%
  filter(mrsa_pre_totvisits_r > 5)

vre_pre_unique_filtered <- vre_pre_unique %>%
  filter(vre_pre_totvisits_r > 5)

rgnb_acq_unique_filtered <- rgnb_acq_unique %>%
  filter(rgnb_acq_totvisits_r > 5)

mrsa_acq_unique_filtered <- mrsa_acq_unique %>%
  filter(mrsa_acq_totvisits_r > 5)

vre_acq_unique_filtered <- vre_acq_unique %>%
  filter(vre_acq_totvisits_r > 5)

# Summary for persistence in rgnb_pre_unique_filtered
summary_persistence_rgnb_pre <- rgnb_pre_unique_filtered %>%
  dplyr::select(Persistence, Clearance) %>%
  tbl_summary(
    type = list(Persistence ~ "categorical",
                Clearance ~ "categorical"),
    statistic = all_categorical() ~ "{n} ({p}%)",
    digits = list(all_categorical() ~ c(0, 1)),
    missing = "no"
  ) %>%
  modify_header(label ~ "**Variable**") %>%
  bold_labels() %>%
  modify_spanning_header(all_stat_cols() ~ "**RGNB**")

# Summary for persistence in mrsa_pre_unique_filtered
summary_persistence_mrsa_pre <- mrsa_pre_unique_filtered %>%
  dplyr::select(Persistence, Clearance) %>%
  tbl_summary(
    type = list(Persistence ~ "categorical",
                Clearance ~ "categorical"),
    statistic = all_categorical() ~ "{n} ({p}%)",
    digits = list(all_categorical() ~ c(0, 1)),
    missing = "no"
  ) %>%
  modify_header(label ~ "**Variable**") %>%
  bold_labels() %>%
  modify_spanning_header(all_stat_cols() ~ "**MRSA**")

# Summary for persistence in vre_pre_unique_filtered
summary_persistence_vre_pre <- vre_pre_unique_filtered %>%
  dplyr::select(Persistence, Clearance) %>%
  tbl_summary(
    type = list(Persistence ~ "categorical",
                Clearance ~ "categorical"),
    statistic = all_categorical() ~ "{n} ({p}%)",
    digits = list(all_categorical() ~ c(0, 1)),
    missing = "no"
  ) %>%
  modify_header(label ~ "**Variable**") %>%
  bold_labels() %>%
  modify_spanning_header(all_stat_cols() ~ "**VRE**")

# Combine summaries for persistence
combined_persistence_summary_pre <- tbl_merge(
  tbls = list(summary_persistence_rgnb_pre, summary_persistence_mrsa_pre, summary_persistence_vre_pre),
  tab_spanner = c("**RGNB**", "**MRSA**", "**VRE**")
) %>%
  as_gt() %>%
  tab_header(
    title = "Persistence and Clearance among NH Residents with Preexisting Colonization"
  ) %>%
  tab_style(
    cell_borders(sides = c("top"), style = "hidden"),
    locations = cells_title()
  )

# View and export the combined persistence table
combined_persistence_summary_pre

# Summary for persistence in rgnb_acq_unique_filtered
summary_persistence_rgnb_acq <- rgnb_acq_unique_filtered %>%
  dplyr::select(Persistence, Clearance) %>%
  tbl_summary(
    type = list(Persistence ~ "categorical",
                Clearance ~ "categorical"),
    statistic = all_categorical() ~ "{n} ({p}%)",
    digits = list(all_categorical() ~ c(0, 1)),
    missing = "no"
  ) %>%
  modify_header(label ~ "**Variable**") %>%
  bold_labels() %>%
  modify_spanning_header(all_stat_cols() ~ "**RGNB**")

# Summary for persistence in mrsa_acq_unique_filtered
summary_persistence_mrsa_acq <- mrsa_acq_unique_filtered %>%
  dplyr::select(Persistence, Clearance) %>%
  tbl_summary(
    type = list(Persistence ~ "categorical",
                Clearance ~ "categorical"),
    statistic = all_categorical() ~ "{n} ({p}%)",
    digits = list(all_categorical() ~ c(0, 1)),
    missing = "no"
  ) %>%
  modify_header(label ~ "**Variable**") %>%
  bold_labels() %>%
  modify_spanning_header(all_stat_cols() ~ "**MRSA**")

# Summary for persistence in vre_acq_unique_filtered
summary_persistence_vre_acq <- vre_acq_unique_filtered %>%
  dplyr::select(Persistence, Clearance) %>%
  tbl_summary(
    type = list(Persistence ~ "categorical",
                Clearance ~ "categorical"),
    statistic = all_categorical() ~ "{n} ({p}%)",
    digits = list(all_categorical() ~ c(0, 1)),
    missing = "no"
  ) %>%
  modify_header(label ~ "**Variable**") %>%
  bold_labels() %>%
  modify_spanning_header(all_stat_cols() ~ "**VRE**")

# Combine summaries for persistence
combined_persistence_summary_acq <- tbl_merge(
  tbls = list(summary_persistence_rgnb_acq, summary_persistence_mrsa_acq, summary_persistence_vre_acq),
  tab_spanner = c("**RGNB**", "**MRSA**", "**VRE**")
) %>%
  as_gt() %>%
  tab_header(
    title = "Persistence and Clearance among NH Residents with Acquired Colonization"
  ) %>%
  tab_style(
    cell_borders(sides = c("top"), style = "hidden"),
    locations = cells_title()
  )

# View and export the combined persistence table
combined_persistence_summary_acq
```


## Revised Version
```{r}
# Load necessary libraries
library(dplyr)
library(gtsummary)
library(gt)
library(broom)

# Function to calculate proportions and 95% CI
calc_prop_ci <- function(data, variable, level) {
  prop <- mean(data[[variable]] == level, na.rm = TRUE) * 100  # Convert to percentage
  ci <- prop.test(sum(data[[variable]] == level, na.rm = TRUE),
                  n = sum(!is.na(data[[variable]])))$conf.int * 100  # Convert to percentage
  return(c(prop, ci))
}

# Helper function to create summary data frame
create_summary_df <- function(data, organism_label) {
  data %>%
    dplyr::select(Persistence, Clearance) %>%
    summarize(
      Organism = organism_label,
      Persistence_Prop = calc_prop_ci(., "Persistence", "Persistent")[1],
      Persistence_Lower_CI = calc_prop_ci(., "Persistence", "Persistent")[2],
      Persistence_Upper_CI = calc_prop_ci(., "Persistence", "Persistent")[3],
      Clearance_Prop = calc_prop_ci(., "Clearance", "Yes")[1],
      Clearance_Lower_CI = calc_prop_ci(., "Clearance", "Yes")[2],
      Clearance_Upper_CI = calc_prop_ci(., "Clearance", "Yes")[3]
    )
}

# Create summary data frames
summary_persistence_rgnb_pre <- create_summary_df(rgnb_pre_unique_filtered, "RGNB")
summary_persistence_mrsa_pre <- create_summary_df(mrsa_pre_unique_filtered, "MRSA")
summary_persistence_vre_pre <- create_summary_df(vre_pre_unique_filtered, "VRE")

# Combine summaries for persistence
combined_persistence_summary_pre <- bind_rows(
  summary_persistence_rgnb_pre,
  summary_persistence_mrsa_pre,
  summary_persistence_vre_pre
)

# Convert to gtsummary table
gt_table_pre <- combined_persistence_summary_pre %>%
  mutate(across(c(Persistence_Prop, Persistence_Lower_CI, Persistence_Upper_CI,
                  Clearance_Prop, Clearance_Lower_CI, Clearance_Upper_CI), 
                ~ sprintf("%.1f%%", .))) %>%
  gt() %>%
  tab_header(
    title = "Persistence and Clearance among NH Residents with Preexisting Colonization"
  ) %>%
  cols_label(
    Organism = "Organism",
    Persistence_Prop = "Persistence (Prop)",
    Persistence_Lower_CI = "Persistence (Lower 95% CI)",
    Persistence_Upper_CI = "Persistence (Upper 95% CI)",
    Clearance_Prop = "Clearance (Prop)",
    Clearance_Lower_CI = "Clearance (Lower 95% CI)",
    Clearance_Upper_CI = "Clearance (Upper 95% CI)"
  ) %>%
  tab_style(
    style = cell_borders(sides = c("top"), style = "hidden"),
    locations = cells_title()
  )

# Print the table
print(gt_table_pre)

# Repeat for acquired colonization
summary_persistence_rgnb_acq <- create_summary_df(rgnb_acq_unique_filtered, "RGNB")
summary_persistence_mrsa_acq <- create_summary_df(mrsa_acq_unique_filtered, "MRSA")
summary_persistence_vre_acq <- create_summary_df(vre_acq_unique_filtered, "VRE")

combined_persistence_summary_acq <- bind_rows(
  summary_persistence_rgnb_acq,
  summary_persistence_mrsa_acq,
  summary_persistence_vre_acq
)

# Convert to gtsummary table
gt_table_acq <- combined_persistence_summary_acq %>%
  mutate(across(c(Persistence_Prop, Persistence_Lower_CI, Persistence_Upper_CI,
                  Clearance_Prop, Clearance_Lower_CI, Clearance_Upper_CI), 
                ~ sprintf("%.1f%%", .))) %>%
  gt() %>%
  tab_header(
    title = "Persistence and Clearance among NH Residents with Acquired Colonization"
  ) %>%
  cols_label(
    Organism = "Organism",
    Persistence_Prop = "Persistence (Prop)",
    Persistence_Lower_CI = "Persistence (Lower 95% CI)",
    Persistence_Upper_CI = "Persistence (Upper 95% CI)",
    Clearance_Prop = "Clearance (Prop)",
    Clearance_Lower_CI = "Clearance (Lower 95% CI)",
    Clearance_Upper_CI = "Clearance (Upper 95% CI)"
  ) %>%
  tab_style(
    style = cell_borders(sides = c("top"), style = "hidden"),
    locations = cells_title()
  )

# Print the table
print(gt_table_acq)
```
```{r}
library(dplyr)
library(gt)
library(epitools)

# Function to calculate proportions and 95% CI in percentage format
calculate_proportion_ci <- function(successes, total) {
  prop <- successes / total * 100
  ci <- prop.test(successes, total)$conf.int * 100
  return(paste0(round(prop, 1), "% (", round(ci[1], 1), "% - ", round(ci[2], 1), "%)"))
}

# Data for Preexisting Colonization
data_preexisting <- data.frame(
  Variable = c("Persistence", "Clearance"),
  RGNB = c(calculate_proportion_ci(64, 88), calculate_proportion_ci(5, 88)),
  MRSA = c(calculate_proportion_ci(33, 44), calculate_proportion_ci(2, 44)),
  VRE = c(calculate_proportion_ci(75, 95), calculate_proportion_ci(10, 95))
)

# Data for New Acquisition
data_new_acquisition <- data.frame(
  Variable = c("Persistence", "Clearance"),
  RGNB = c(calculate_proportion_ci(20, 33), calculate_proportion_ci(3, 33)),
  MRSA = c(calculate_proportion_ci(10, 12), calculate_proportion_ci(0, 12)),
  VRE = c(calculate_proportion_ci(10, 20), calculate_proportion_ci(1, 20))
)

# Combine data frames into one table
combined_data <- bind_rows(
  data_preexisting %>% mutate(Category = "Preexisting Colonization"),
  data_new_acquisition %>% mutate(Category = "New Acquisition")
)

# Format the table using gt
combined_table <- combined_data %>%
  select(Category, Variable, RGNB, MRSA, VRE) %>%
  gt() %>%
  tab_header(
    title = "Persistence and Clearance among NH Residents without Intervention",
    subtitle = "Proportions and 95% Confidence Intervals"
  ) %>%
  cols_label(
    Category = "Category",
    Variable = "Variable",
    RGNB = "RGNB",
    MRSA = "MRSA",
    VRE = "VRE"
  ) %>%
  tab_spanner(
    label = "Preexisting Colonization",
    columns = c(RGNB, MRSA, VRE)
  ) %>%
  fmt_missing(
    columns = everything(),
    missing_text = "0.0% (0.0% - 0.0%)"
  )

# Display the table
print(combined_table)
```


## Type A plot
```{r}
# Load necessary libraries
library(ggplot2)
library(dplyr)

# Data for pre-existing colonization
data_pre <- data.frame(
  organism = rep(c("RGNB", "MRSA", "VRE"), each = 2),
  measure = rep(c("Persistence", "Clearance"), times = 3),
  proportion = c(72.7, 9.4, 75.0, 12.1, 78.9, 17.3),
  lower_ci = c(62.0, 3.9, 59.4, 4.0, 69.1, 9.9),
  upper_ci = c(81.4, 19.9, 86.3, 29.1, 86.4, 28.2)
)

# Data for new acquisition colonization
data_acq <- data.frame(
  organism = rep(c("RGNB", "MRSA", "VRE"), each = 2),
  measure = rep(c("Persistence", "Clearance"), times = 3),
  proportion = c(60.6, 30.0, 83.3, 0.0, 50.0, 20.0),
  lower_ci = c(42.2, 12.8, 50.9, 0.0, 29.9, 3.5),
  upper_ci = c(76.6, 54.3, 97.1, 34.5, 70.1, 55.8)
)

# Function to plot proportions
plot_proportions <- function(data, title) {
  ggplot(data, aes(x = organism, y = proportion, color = measure)) +
    geom_point(position = position_dodge(width = 0.5)) +
    geom_errorbar(aes(ymin = lower_ci, ymax = upper_ci), width = 0.2, position = position_dodge(width = 0.5)) +
    labs(
      title = title,
      x = "Organism",
      y = "Proportion (%)"
    ) +
    theme_minimal() +
    scale_color_manual(values = c("Persistence" = "blue", "Clearance" = "red"))
}

# Plot for pre-existing
plot_pre <- plot_proportions(data_pre, "Proportions for Pre-existing Colonization")

# Plot for new acquisition
plot_acq <- plot_proportions(data_acq, "Proportions for New-Acquisition Colonization")

# Print the plots
print(plot_pre)
print(plot_acq)
```
## Type B plots
```{r}
# Load necessary libraries
library(ggplot2)
library(dplyr)

# Data for persistence
data_persistence <- data.frame(
  organism = rep(c("RGNB", "MRSA", "VRE"), times = 2),
  measure = rep(c("Pre-existing Colonization", "New Acquisition"), each = 3),
  proportion = c(72.7, 75.0, 78.9, 60.6, 83.3, 50.0),
  lower_ci = c(62.0, 59.4, 69.1, 42.2, 50.9, 29.9),
  upper_ci = c(81.4, 86.3, 86.4, 76.6, 97.1, 70.1)
)

# Data for clearance
data_clearance <- data.frame(
  organism = rep(c("RGNB", "MRSA", "VRE"), times = 2),
  measure = rep(c("Pre-existing Colonization", "New Acquisition"), each = 3),
  proportion = c(5.7, 4.5, 10.5, 9.1, 0.0, 5.0),
  lower_ci = c(2.1, 0.8, 5.4, 2.4, 0.0, 0.3),
  upper_ci = c(13.4, 16.7, 18.9, 25.5, 30.1, 26.9)
)

# Function to plot proportions with custom y-axis limits
plot_proportions <- function(data, title, y_max = 100) {
  ggplot(data, aes(x = organism, y = proportion, color = measure)) +
    geom_point(position = position_dodge(width = 0.5)) +
    geom_errorbar(aes(ymin = lower_ci, ymax = upper_ci), width = 0.2, position = position_dodge(width = 0.5)) +
    labs(
      title = title,
      x = "Organism",
      y = "Proportion (95% CI)"
    ) +
    theme_minimal() +
    scale_color_manual(values = c("Pre-existing Colonization" = "blue", "New Acquisition" = "orange"))
}

# Plot for persistence
plot_persistence <- plot_proportions(data_persistence, "Proportions for Persistence")

# Plot for clearance
plot_clearance <- plot_proportions(data_clearance, "Proportions for Clearance")

# Print the plots
print(plot_persistence)
print(plot_clearance)

# Save the plots to files
ggsave("persistence_plot.png", plot = plot_persistence, width = 8, height = 6, dpi = 300)
ggsave("clearance_plot.png", plot = plot_clearance, width = 8, height = 6, dpi = 300)
```
```{r}
# Load necessary libraries
library(ggplot2)
library(dplyr)

# Data for persistence
data_persistence <- data.frame(
  organism = rep(c("RGNB", "MRSA", "VRE"), times = 2),
  measure = factor(rep(c("Pre-existing Colonization", "New Acquisition"), each = 3), levels = c("Pre-existing Colonization", "New Acquisition")),
  proportion = c(72.7, 75.0, 78.9, 60.6, 83.3, 50.0),
  lower_ci = c(62.0, 59.4, 69.1, 42.2, 50.9, 29.9),
  upper_ci = c(81.4, 86.3, 86.4, 76.6, 97.1, 70.1)
)

# Data for clearance
data_clearance <- data.frame(
  organism = rep(c("RGNB", "MRSA", "VRE"), times = 2),
  measure = factor(rep(c("Pre-existing Colonization", "New Acquisition"), each = 3), levels = c("Pre-existing Colonization", "New Acquisition")),
  proportion = c(35.7, 50.0, 71.4, 50.0 , 0.0, 50.0),
  lower_ci = c(10.6, 1.0, 47.8, 10.0, 0.0, -19.3),
  upper_ci = c(60.8, 99.0, 95.1, 90.0, 0.0, 119.3)
)

# Function to plot proportions with custom y-axis limits
plot_proportions <- function(data, title, y_max = 100) {
  ggplot(data, aes(x = organism, y = proportion, fill = measure)) +
    geom_bar(stat = "identity", position = position_dodge(width = 0.9), color = "black", width = 0.7) +
    geom_errorbar(aes(ymin = lower_ci, ymax = upper_ci), width = 0.2, position = position_dodge(width = 0.9)) +
    geom_text(aes(label = round(proportion, 1), y = proportion / 2), 
              position = position_dodge(width = 0.9), 
              vjust = 0.5, size = 3, color = "white") +
    labs(
      title = title,
      x = "Organism",
      y = "Proportion (95% CI)"
    ) +
    theme_minimal() +
    scale_fill_manual(values = c("Pre-existing Colonization" = "blue", "New Acquisition" = "orange"))
}

# Plot for persistence
plot_persistence <- plot_proportions(data_persistence, "Proportions of MDRO Persistence", y_max = 100)

# Plot for clearance
plot_clearance <- plot_proportions(data_clearance, "Proportions of MDRO Clearance", y_max = 100)

# Print the plots
print(plot_persistence)
print(plot_clearance)

# Save the plots to files
ggsave("persistence_bar.png", plot = plot_persistence, width = 8, height = 6, dpi = 300)
ggsave("clearance_bar.png", plot = plot_clearance, width = 8, height = 6, dpi = 300)
```


## Type C plots
```{r}
# Load necessary libraries
library(ggplot2)
library(dplyr)
library(ggpubr)

# Data for persistence
data_persistence <- data.frame(
  organism = rep(c("RGNB", "MRSA", "VRE"), times = 2),
  measure = rep(c("Pre-existing", "New-acquisition"), each = 3),
  proportion = c(72.7, 75.0, 78.9, 60.6, 83.3, 50.0),
  lower_ci = c(62.0, 59.4, 69.1, 42.2, 50.9, 29.9),
  upper_ci = c(81.4, 86.3, 86.4, 76.6, 97.1, 70.1)
)

# Data for clearance
data_clearance <- data.frame(
  organism = rep(c("RGNB", "MRSA", "VRE"), times = 2),
  measure = rep(c("Pre-existing", "New-acquisition"), each = 3),
  proportion = c(9.4, 12.1, 17.3, 30.0, 0.0, 20.0),
  lower_ci = c(3.9, 4.0, 9.9, 12.8, 0.0, 3.5),
  upper_ci = c(19.9, 29.1, 28.2, 54.3, 34.5, 55.8)
)

# Function to calculate p-values using chi-squared test
calc_p_values <- function(data) {
  p_values <- data %>%
    group_by(organism) %>%
    summarize(p_value = chisq.test(matrix(c(
      proportion[measure == "Pre-existing"], 100 - proportion[measure == "Pre-existing"],
      proportion[measure == "New-acquisition"], 100 - proportion[measure == "New-acquisition"]
    ), nrow = 2))$p.value)
  return(p_values)
}

# Calculate p-values for persistence and clearance
p_values_persistence <- calc_p_values(data_persistence)
p_values_clearance <- calc_p_values(data_clearance)

# Function to plot proportions and add p-values
plot_proportions <- function(data, title, p_values) {
  ggplot(data, aes(x = organism, y = proportion, color = measure)) +
    geom_point(position = position_dodge(width = 0.5)) +
    geom_errorbar(aes(ymin = lower_ci, ymax = upper_ci), width = 0.2, position = position_dodge(width = 0.5)) +
    labs(
      title = title,
      x = "Organism",
      y = "Proportion (%)"
    ) +
    theme_minimal() +
    scale_color_manual(values = c("Pre-existing" = "blue", "New-acquisition" = "red")) +
    geom_text(data = p_values, aes(x = organism, y = 105, label = sprintf("p = %.3f", p_value)), color = "black")
}

# Plot for persistence
plot_persistence <- plot_proportions(data_persistence, "Proportions for Persistence", p_values_persistence)

# Plot for clearance
plot_clearance <- plot_proportions(data_clearance, "Proportions for Clearance", p_values_clearance)

# Print the plots
print(plot_persistence)
print(plot_clearance)

# Save the plots to files
ggsave("persistence_plot.png", plot = plot_persistence, width = 8, height = 6, dpi = 300)
ggsave("clearance_plot.png", plot = plot_clearance, width = 8, height = 6, dpi = 300)

```
## Type D plots
```{r}
# Load necessary libraries
library(ggplot2)
library(dplyr)

# Data for persistence
data_persistence <- data.frame(
  organism = rep(c("RGNB", "MRSA", "VRE"), times = 2),
  measure = rep(c("Pre-existing Colonization", "New Acquisition"), each = 3),
  proportion = c(72.7, 75.0, 78.9, 60.6, 83.3, 50.0),
  lower_ci = c(62.0, 59.4, 69.1, 42.2, 50.9, 29.9),
  upper_ci = c(81.4, 86.3, 86.4, 76.6, 97.1, 70.1)
)

# Data for clearance
data_clearance <- data.frame(
  organism = rep(c("RGNB", "MRSA", "VRE"), times = 2),
  measure = rep(c("Pre-existing Colonization", "New Acquisition"), each = 3),
  proportion = c(5.7, 4.5, 10.5, 9.1, 0.0, 5.0),
  lower_ci = c(2.1, 0.8, 5.4, 2.4, 0.0, 0.3),
  upper_ci = c(13.4, 16.7, 18.9, 25.5, 30.1, 26.9)
)

# Function to plot proportions with custom y-axis limits
plot_proportions <- function(data, title, y_max = 100) {
  ggplot(data, aes(x = organism, y = proportion, color = measure)) +
    geom_point(position = position_dodge(width = 0.5)) +
    geom_errorbar(aes(ymin = lower_ci, ymax = upper_ci), width = 0.2, position = position_dodge(width = 0.5)) +
    labs(
      title = title,
      x = "Organism",
      y = "Proportion (95% CI)"
    ) +
    theme_minimal() +
    scale_color_manual(values = c("Pre-existing Colonization" = "blue", "New Acquisition" = "orange")) +
    scale_y_continuous(limits = c(0, y_max))
}

# Plot for persistence
plot_persistence <- plot_proportions(data_persistence, "Proportions for Persistence", y_max = 100)

# Plot for clearance
plot_clearance <- plot_proportions(data_clearance, "Proportions for Clearance", y_max = 100)

# Print the plots
print(plot_persistence)
print(plot_clearance)

# Save the plots to files
ggsave("persistence_plot_100.png", plot = plot_persistence, width = 8, height = 6, dpi = 300)
ggsave("clearance_plot_100.png", plot = plot_clearance, width = 8, height = 6, dpi = 300)
```


## Add chi-square p-value in persistence among organisms
```{r}
# Define the contingency tables
table1 <- matrix(c(24, 11, 20, 64, 33, 75), nrow = 2, byrow = TRUE)
table2 <- matrix(c(58, 29, 62, 6, 4, 13), nrow = 2, byrow = TRUE)
table3 <- matrix(c(13, 2, 10, 20, 10, 10), nrow = 2, byrow = TRUE)
table4 <- matrix(c(14, 10, 8, 6, 0, 2), nrow = 2, byrow = TRUE)

# Function to decide which test to use and perform it
perform_test <- function(table) {
  # Calculate expected frequencies
  expected_frequencies <- chisq.test(table)$expected
  
  # Decide which test to use
  if (all(expected_frequencies >= 5)) {
    # Use chi-square test
    test_result <- chisq.test(table)
  } else {
    # Use Fisher's exact test
    test_result <- fisher.test(table)
  }
  
  return(test_result)
}

# Perform tests on all tables
test_result1 <- perform_test(table1)
test_result2 <- perform_test(table2)
test_result3 <- perform_test(table3)
test_result4 <- perform_test(table4)

# Print results
print("Preexisting Colonization - Persistence")
print(test_result1)

print("Preexisting Colonization - Clearance")
print(test_result2)

print("New Acquisition - Persistence")
print(test_result3)

print("New Acquisition - Clearance")
print(test_result4)

```

# Table 2
## Duration (Colonization) with persistence
```{r}
# Function to summarize duration (with persistence)
summarize_duration <- function(data, label) {
  data %>%
    filter(Persistence == "Persistent") %>%
    summarize(
      N = n(),
      Mean = mean(Duration, na.rm = TRUE),
      SD = sd(Duration, na.rm = TRUE),
      Median = median(Duration, na.rm = TRUE),
      Min = min(Duration, na.rm = TRUE),
      Max = max(Duration, na.rm = TRUE)
    ) %>%
    mutate(Variable = label) %>%
    select(Variable, everything())
}

# Function to summarize duration (cleared)
summarize_duration_c <- function(data, label) {
  data %>%
    filter(Clearance == "Yes") %>%
    summarize(
      N = n(),
      Mean = mean(Duration, na.rm = TRUE),
      SD = sd(Duration, na.rm = TRUE),
      Median = median(Duration, na.rm = TRUE),
      Min = min(Duration, na.rm = TRUE),
      Max = max(Duration, na.rm = TRUE)
    ) %>%
    mutate(Variable = label) %>%
    select(Variable, everything())
}
```

```{r}
# Ensure the Duration column is numeric
rgnb_pre_unique$Duration <- as.numeric(rgnb_pre_unique$Duration)
mrsa_pre_unique$Duration <- as.numeric(mrsa_pre_unique$Duration)
vre_pre_unique$Duration <- as.numeric(vre_pre_unique$Duration)

# Summarize duration for each dataset
summary_duration_rgnb_c <- summarize_duration(rgnb_pre_unique, "RGNB")
summary_duration_mrsa_c <- summarize_duration(mrsa_pre_unique, "MRSA")
summary_duration_vre_c <- summarize_duration(vre_pre_unique, "VRE")

# Combine the summaries into one dataframe
combined_duration_summary_c <- bind_rows(summary_duration_rgnb_c, summary_duration_mrsa_c, summary_duration_vre_c)

# Convert to gt table
duration_gt_table_c <- combined_duration_summary_c %>%
  gt() %>%
  tab_header(
    title = "Duration of Persistence in Preexisting Colonization - Persistent Patients"
  ) %>%
  fmt_number(
    columns = c(Mean, SD, Median, Min, Max),
    decimals = 2
  ) %>%
    tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_body(columns = Variable)
  ) %>%
  tab_options(
    column_labels.font.weight = "bold"
  )
  
# View the combined duration table
duration_gt_table_c
```
## Duration (Colonization) with cleared
```{r}
# Ensure the Duration column is numeric
rgnb_pre_unique$Duration <- as.numeric(rgnb_pre_unique$Duration)
mrsa_pre_unique$Duration <- as.numeric(mrsa_pre_unique$Duration)
vre_pre_unique$Duration <- as.numeric(vre_pre_unique$Duration)

rgnb_pre_unique_filtered <- rgnb_pre_unique %>%
  filter(rgnb_pre_totvisits_r > 5)

mrsa_pre_unique_filtered <- mrsa_pre_unique %>%
  filter(mrsa_pre_totvisits_r > 5)

vre_pre_unique_filtered <- vre_pre_unique %>%
  filter(vre_pre_totvisits_r > 5)

# Summarize duration for each dataset
summary_duration_rgnb_c_c <- summarize_duration_c(rgnb_pre_unique_filtered, "RGNB")
summary_duration_mrsa_c_c <- summarize_duration_c(mrsa_pre_unique_filtered, "MRSA")
summary_duration_vre_c_c <- summarize_duration_c(vre_pre_unique_filtered, "VRE")

# Combine the summaries into one dataframe
combined_duration_summary_c_c <- bind_rows(summary_duration_rgnb_c_c, summary_duration_mrsa_c_c, summary_duration_vre_c_c)

# Convert to gt table
duration_gt_table_c_c <- combined_duration_summary_c_c %>%
  gt() %>%
  tab_header(
    title = "Duration of Persistence in Preexisting Colonization - Cleared Patients"
  ) %>%
  fmt_number(
    columns = c(Mean, SD, Median, Min, Max),
    decimals = 2
  ) %>%
    tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_body(columns = Variable)
  ) %>%
  tab_options(
    column_labels.font.weight = "bold"
  )

# View the combined duration table
duration_gt_table_c_c
```
## Duration (NewAcquisition) with persistence
```{r}
# Ensure the Duration column is numeric
rgnb_acq_unique$Duration <- as.numeric(rgnb_acq_unique$Duration)
mrsa_acq_unique$Duration <- as.numeric(mrsa_acq_unique$Duration)
vre_acq_unique$Duration <- as.numeric(vre_acq_unique$Duration)

rgnb_acq_unique_filtered <- rgnb_acq_unique %>%
  filter(rgnb_acq_totvisits_r > 5)

mrsa_acq_unique_filtered <- mrsa_acq_unique %>%
  filter(mrsa_acq_totvisits_r > 5)

vre_acq_unique_filtered <- vre_acq_unique %>%
  filter(vre_acq_totvisits_r > 5)

# Summarize duration for each dataset
summary_duration_rgnb_a <- summarize_duration(rgnb_acq_unique, "RGNB")
summary_duration_mrsa_a <- summarize_duration(mrsa_acq_unique, "MRSA")
summary_duration_vre_a <- summarize_duration(vre_acq_unique, "VRE")

# Combine the summaries into one dataframe
combined_duration_summary_a <- bind_rows(summary_duration_rgnb_a, summary_duration_mrsa_a, summary_duration_vre_a)

# Convert to gt table
duration_gt_table_a <- combined_duration_summary_a %>%
  gt() %>%
  tab_header(
    title = "Duration of Persistence in New Acquisition - Persistent Patients"
  ) %>%
  fmt_number(
    columns = c(Mean, SD, Median, Min, Max),
    decimals = 2
  ) %>%
    tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_body(columns = Variable)
  ) %>%
  tab_options(
    column_labels.font.weight = "bold"
  )
  
# View the combined duration table
duration_gt_table_a
```
## Duration (NewAcquisition) with cleared
```{r}
# Summarize duration for each dataset
summary_duration_rgnb_a_c <- summarize_duration_c(rgnb_acq_unique_filtered, "RGNB")
summary_duration_mrsa_a_c <- summarize_duration_c(mrsa_acq_unique_filtered, "MRSA")
summary_duration_vre_a_c <- summarize_duration_c(vre_acq_unique_filtered, "VRE")

# Combine the summaries into one dataframe
combined_duration_summary_a_c <- bind_rows(summary_duration_rgnb_a_c, summary_duration_mrsa_a_c, summary_duration_vre_a_c)

# Convert to gt table
duration_gt_table_a_c <- combined_duration_summary_a_c %>%
  gt() %>%
  tab_header(
    title = "Duration of Persistence in New Acquisition - Cleared Patients"
  ) %>%
  fmt_number(
    columns = c(Mean, SD, Median, Min, Max),
    decimals = 2
  ) %>%
    tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_body(columns = Variable)
  ) %>%
  tab_options(
    column_labels.font.weight = "bold"
  )

# View the combined duration table
duration_gt_table_a_c
```
# Figure 1.Compare duration in perexisting vs. new acquisition (persistence)
```{r}
# Combine data frames for pre-existing and new acquisition
combined_duration_rgnb <- bind_rows(
  rgnb_pre_unique %>% mutate(Colonization = "Pre-existing"),
  rgnb_acq_unique %>% mutate(Colonization = "New Acquisition"),
  .id = "id")

# Combine data frames for pre-existing and new acquisition
combined_duration_mrsa <- bind_rows(
  mrsa_pre_unique %>% mutate(Colonization = "Pre-existing"),
 mrsa_acq_unique %>% mutate(Colonization = "New Acquisition"),
  .id = "id")

# Combine data frames for pre-existing and new acquisition
combined_duration_vre <- bind_rows(
  vre_pre_unique %>% mutate(Colonization = "Pre-existing"),
  vre_acq_unique %>% mutate(Colonization = "New Acquisition"),
  .id = "id")

# Combine all of them
combined_duration <- bind_rows(combined_duration_rgnb, combined_duration_mrsa, combined_duration_vre, .id = "organism")
combined_duration <- combined_duration %>%
  mutate(organism = case_when(
    organism == 1 ~ "RGNB",
    organism == 2 ~ "MRSA",
    organism == 3 ~ "VRE"
  ))

# Check the distribution
hist(combined_duration_rgnb$Duration, main="Histogram of Duration"
, xlab="Duration")
qqnorm(combined_duration_rgnb$Duration, main = "Q-Q Plot of Duration")
qqline(combined_duration_rgnb$Duration) # Not normally distribution
shapiro.test(combined_duration_rgnb$Duration)

hist(combined_duration_mrsa$Duration, main="Histogram of Duration"
, xlab="Duration")
qqnorm(combined_duration_mrsa$Duration, main = "Q-Q Plot of Duration")
qqline(combined_duration_mrsa$Duration) # Not normally distribution
shapiro.test(combined_duration_mrsa$Duration)

hist(combined_duration_vre$Duration, main="Histogram of Duration"
, xlab="Duration")
qqnorm(combined_duration_vre$Duration, main = "Q-Q Plot of Duration")
qqline(combined_duration_vre$Duration) # Not normally distribution
shapiro.test(combined_duration_vre$Duration)

# Create boxplot with p values
p1 <- ggboxplot(combined_duration, 
                x = "Colonization", y = "Duration",
                color = "Colonization", palette = "jco",
                add = "jitter",
                facet.by = "organism") +
  stat_compare_means(method = "wilcox.test") + 
  labs(y = "Duration of persistence")  # Add y-axis label
p1

ggsave("boxplot_1.png", plot = p1, width = 8, height = 6, dpi = 600)
```

```{r}
library(dplyr)
library(gtsummary)
library(ggplot2)
library(ggpubr)

# Ensure all data frames have the same columns and types
common_columns <- intersect(intersect(names(rgnb_pre_unique_clearance), names(mrsa_pre_unique_clearance)), names(vre_pre_unique_clearance))

# Convert cogsum_bl to numeric type in all data frames
convert_to_numeric <- function(df) {
  df %>% 
    select(all_of(common_columns)) %>%
    mutate(cogsum_bl = as.numeric(cogsum_bl))
}

rgnb_pre_unique_clearance <- convert_to_numeric(rgnb_pre_unique_clearance)
mrsa_pre_unique_clearance <- convert_to_numeric(mrsa_pre_unique_clearance)
vre_pre_unique_clearance <- convert_to_numeric(vre_pre_unique_clearance)
rgnb_acq_unique_clearance <- convert_to_numeric(rgnb_acq_unique_clearance)
mrsa_acq_unique_clearance <- convert_to_numeric(mrsa_acq_unique_clearance)
vre_acq_unique_clearance <- convert_to_numeric(vre_acq_unique_clearance)

# Combine data frames for pre-existing and new acquisition for each organism
combined_duration_rgnb_clearance <- bind_rows(
  rgnb_pre_unique_clearance %>% mutate(Colonization = "Pre-existing"),
  rgnb_acq_unique_clearance %>% mutate(Colonization = "New Acquisition"),
  .id = "id")

combined_duration_mrsa_clearance <- bind_rows(
  mrsa_pre_unique_clearance %>% mutate(Colonization = "Pre-existing"),
  mrsa_acq_unique_clearance %>% mutate(Colonization = "New Acquisition"),
  .id = "id")

combined_duration_vre_clearance <- bind_rows(
  vre_pre_unique_clearance %>% mutate(Colonization = "Pre-existing"),
  vre_acq_unique_clearance %>% mutate(Colonization = "New Acquisition"),
  .id = "id")

# Combine all of them into one dataset
combined_duration_clearance <- bind_rows(combined_duration_rgnb_clearance, combined_duration_mrsa_clearance, combined_duration_vre_clearance, .id = "organism")
combined_duration_clearance <- combined_duration_clearance %>%
  mutate(organism = case_when(
    organism == "1" ~ "RGNB",
    organism == "2" ~ "MRSA",
    organism == "3" ~ "VRE"
  ))

# Create boxplot with p values
p2 <- ggboxplot(combined_duration_clearance, 
                x = "Colonization", y = "Duration",
                color = "Colonization", palette = "jco",
                add = "jitter",
                facet.by = "organism") +
  stat_compare_means(method = "wilcox.test") + 
  labs(y = "Duration of persistence")  # Add y-axis label

p2

# Save the boxplot
ggsave("boxplot_2_clearance.png", plot = p2, width = 8, height = 6, dpi = 600)


```

# Figure 2. Compare duration with clearance of pre-existing colonization among species
```{r}
#library(dplyr)
#library(ggpubr)
#library(ggplot2)

# Perorm pairwise comparisons
#compare_means(Duration ~ organism,  data = combined_duration_2)

# Change Colonization Order
#combined_duration_2$Colonization = factor(combined_duration_2$Colonization,
                                          #levels = c("Pre-existing", "New Acquisition"))

# Create boxplot with p values
#my_comparisons <- list( c("RGNB", "MRSA"), c("MRSA", "VRE"), c("RGNB", "VRE") )
#p2 <- ggboxplot(combined_duration_2, 
                #x = "organism", 
                #y = "duration",
                #color = "organism", 
                #palette = "jco",
                #add = "jitter",
                #facet.by = "Colonization") +
#  stat_compare_means(comparisons = my_comparisons)+ # Add pairwise comparisons p-value
  #stat_compare_means(label.y = 100)     # Add global p-value + 
  #labs(y = "Duration of persistence")  # Add y-axis label

# Display the plot
#print(p2)

# Save the plot as a PNG file
#ggsave("boxplot_3.png", plot = p2, width = 8, height = 6, dpi = 600)
```
```{r}
# Convert variables as necessary
rgnb_pre_unique$Hands_only <- as.character(rgnb_pre_unique$Hands_only)
rgnb_pre_unique$Age <- as.numeric(rgnb_pre_unique$age_bl)
rgnb_pre_unique$cogsum_bl <- as.numeric(rgnb_pre_unique$cogsum_bl)
var_label(rgnb_pre_unique$cogsum_bl) <- "Cognitive Sum Score"

# Univariate logistic regression for odds ratios
rgnb_uni_or <- rgnb_pre_unique %>%
  dplyr::select(Age,
                Sex,
                Race,
                gnbr_p_cat,
                gnbr_env_cat,
                antibiotic_bl,
                deviceuse, 
                openwound_final,
                cogsum_bl,
                horgnb_bl,
                psmscore_cat,
                charlson_cat,
                hospstay_cat,
                Persistence,
                Diabetes,
                ccsdemen_bl,
                Hands_only
) %>%
  tbl_uvregression(
    method = glm,
    y = Persistence,
    method.args = list(family = binomial), # logistic regression
    exponentiate = TRUE # to display OR instead of log(OR)
  ) %>%
  add_global_p() %>% # adds p-values for the overall test
  modify_header(label ~ "**Characteristic**") %>% # Label the variables column
  modify_column_header(
    estimate ~ "**Unadjusted OR (95% CI)**",
    p.value ~ "**P value**"
  ) %>%
  bold_labels() %>%
  as_gt() %>%
  tab_header(
    title = "Table 2. Association between Participant Characteristics and Baseline MDRO Colonization or New MDRO Acquisitions"
  ) %>%
  tab_style(cell_borders(
    sides = c("top"), style = "hidden"),
    locations = cells_title()
  )

# Display the table
rgnb_uni_or
```

# Table 3: Univariate Analysis for Persistence Outcome
## RGNB Univariate_Colonization
```{r}
#| label: RGNB univariate analysis_colonization
# Colonization RGNB Univariate Table (Crude)
rgnb_pre_unique$Hands_only <- as.character(rgnb_pre_unique$Hands_only)

rgnb_pre_unique$Age <- as.numeric(rgnb_pre_unique$age_bl)
rgnb_pre_unique$cogsum_bl <- as.numeric(rgnb_pre_unique$cogsum_bl)
var_label(rgnb_pre_unique$cogsum_bl) <- "Cognitive Sum Score"

rgnb_uni_c <- rgnb_pre_unique %>%
  dplyr::select(Age,
                Sex,
                Race,
                gnbr_p_cat,
                gnbr_env_cat,
                antibiotic_bl,
                deviceuse, 
                openwound_final,
                cogsum_bl,
                horgnb_bl,
                psmscore_cat,
                charlson_cat,
                hospstay_cat,
                Persistence,
                Diabetes,
                ccsdemen_bl,
                Hands_only
) %>%
  tbl_summary(by = Persistence, # stratify by persistence
              statistic = list(all_continuous() ~ "{mean} ({sd})", #use mean and standard deviation for continuous variables
                               all_categorical() ~ "{n} ({p}%)"), #use count and percentage for categorical variables
              digits = list(all_categorical() ~ c(0, 2), #adds no decimal places to counts and one decimal to percentages
                            all_continuous() ~ 1), #adds one decimal place to mean and sd values
              missing = "no"
              #missing_text = "Missing (n)"
              ) %>%
  add_p(pvalue_fun = ~style_pvalue(.x, digits = 2)) %>% #adds p value
  add_overall() %>% #adds column with non-stratified summary statistics
  modify_header(label ~ "**Variable**") %>% #the asterisks bold the label
  modify_spanning_header(c("stat_1", "stat_2") ~ "**Persistence**") %>%
  bold_labels() %>%
  as_gt() %>% 
  tab_header(
    title = "Table 3. Variables associated with RGNB persistence in perexisting colonization patient"
    ) %>% 
  tab_style(cell_borders(
    sides = c("top"), style = "hidden"),
    locations = cells_title())

# View the table
rgnb_uni_c
```
# RGNB OR
```{r}
# Filter out variables with only one unique level
categorical_vars_rgnb <- c("Sex", "gnbr_p_cat", "gnbr_env_cat", "antibiotic_bl", 
                      "deviceuse", "openwound_final", "horgnb_bl", "psmscore_cat", 
                      "charlson_cat", "hospstay_cat", "Persistence", "Diabetes", 
                      "ccsdemen_bl", "Hands_only")

# Filter for categorical variables with at least two levels
valid_categorical_vars_rgnb <- categorical_vars_rgnb[sapply(rgnb_pre_unique[categorical_vars_rgnb], function(x) length(unique(x)) > 1)]

# Select only valid variables for analysis
rgnb_uni_or <- rgnb_pre_unique %>%
  dplyr::select(Age, cogsum_bl, all_of(valid_categorical_vars)) %>%
  tbl_uvregression(
    method = glm,
    y = Persistence,
    method.args = list(family = binomial), # logistic regression
    exponentiate = TRUE # to display OR instead of log(OR)
  ) %>%
  add_global_p() %>% # adds p-values for the overall test
  modify_header(
    label ~ "**Characteristic**",
    estimate ~ "**Unadjusted OR (95% CI)**",
    p.value ~ "**P value**"
  ) %>%
  bold_labels() %>%
  as_gt() %>%
  tab_header(
    title = "Table 1. Association between Participant Characteristics and Baseline MDRO Colonization or New MDRO Acquisitions"
  ) %>%
  tab_style(cell_borders(
    sides = c("top"), style = "hidden"),
    locations = cells_title()
  )

# Display the table
rgnb_uni_or

```
# MRSA OR
```{r}
# Filter out variables with only one unique level
categorical_vars_mrsa <- c("Sex", "mrsa_p_cat", "mrsa_env_cat", "antibiotic_bl", 
                      "deviceuse", "openwound_final", "homrsa_bl", "psmscore_cat", 
                      "charlson_cat", "hospstay_cat", "Persistence", "Diabetes", 
                      "ccschf_bl", "Hands_only")

mrsa_pre_unique$age_bl <- as.numeric(mrsa_pre_unique$age_bl)
mrsa_pre_unique$cogsum_bl <- as.numeric(mrsa_pre_unique$cogsum_bl)

# Filter for categorical variables with at least two levels
valid_categorical_vars_mrsa <- categorical_vars_mrsa[sapply(mrsa_pre_unique[categorical_vars_mrsa], function(x) length(unique(x)) > 1)]

# Select only valid variables for analysis
mrsa_uni_or <- mrsa_pre_unique %>%
  dplyr::select(age_bl, cogsum_bl, all_of(valid_categorical_vars_mrsa)) %>%
  tbl_uvregression(
    method = glm,
    y = Persistence,
    method.args = list(family = binomial), # logistic regression
    exponentiate = TRUE # to display OR instead of log(OR)
  ) %>%
  add_global_p() %>% # adds p-values for the overall test
  modify_header(
    label ~ "**Characteristic**",
    estimate ~ "**Unadjusted OR (95% CI)**",
    p.value ~ "**P value**"
  ) %>%
  bold_labels() %>%
  as_gt() %>%
  tab_header(
    title = "Table 1. Association between Participant Characteristics and Baseline MDRO Colonization or New MDRO Acquisitions"
  ) %>%
  tab_style(cell_borders(
    sides = c("top"), style = "hidden"),
    locations = cells_title()
  )

# Display the table
mrsa_uni_or

```
# VRE OR
```{r}
# Filter out variables with only one unique level
categorical_vars_vre <- c("Sex", "vre_p_cat", "vre_env_cat", "antibiotic_bl", 
                      "deviceuse", "openwound_final", "hovre_bl", "psmscore_cat", 
                      "charlson_cat", "hospstay_cat", "Persistence", "Diabetes", 
                      "ccscopd_bl", "Hands_only")
vre_pre_unique$age_bl <- as.numeric(vre_pre_unique$age_bl)
vre_pre_unique$cogsum_bl <- as.numeric(vre_pre_unique$cogsum_bl)

# Filter for categorical variables with at least two levels
valid_categorical_vars_vre <- categorical_vars_vre[sapply(vre_pre_unique[categorical_vars_vre], function(x) length(unique(x)) > 1)]

# Select only valid variables for analysis
vre_uni_or <- vre_pre_unique %>%
  dplyr::select(age_bl, cogsum_bl, all_of(valid_categorical_vars_vre)) %>%
  tbl_uvregression(
    method = glm,
    y = Persistence,
    method.args = list(family = binomial), # logistic regression
    exponentiate = TRUE # to display OR instead of log(OR)
  ) %>%
  add_global_p() %>% # adds p-values for the overall test
  modify_header(
    label ~ "**Characteristic**",
    estimate ~ "**Unadjusted OR (95% CI)**",
    p.value ~ "**P value**"
  ) %>%
  bold_labels() %>%
  as_gt() %>%
  tab_header(
    title = "Table 1. Association between Participant Characteristics and Baseline MDRO Colonization or New MDRO Acquisitions_VRE"
  ) %>%
  tab_style(cell_borders(
    sides = c("top"), style = "hidden"),
    locations = cells_title()
  )

# Display the table
vre_uni_or

```

## MRSA Univariate_Colonization
```{r}
#| label: MRSA univariate analysis_colonization
# Colonization MRSA Univariate Table (Crude)
mrsa_pre_unique$Hands_only <- as.character(mrsa_pre_unique$Hands_only)

mrsa_pre_unique$Age <- as.numeric(mrsa_pre_unique$age_bl)
mrsa_pre_unique$cogsum_bl <- as.numeric(mrsa_pre_unique$cogsum_bl)
var_label(mrsa_pre_unique$cogsum_bl) <- "Cognitive Sum Score"

mrsa_uni_c <- mrsa_pre_unique %>%
  dplyr::select(Age,
                Sex,
                Race,
                mrsa_p_cat,
                mrsa_env_cat,
                antibiotic_bl,
                deviceuse, 
                openwound_final,
                cogsum_bl,
                homrsa_bl,
                psmscore_cat,
                charlson_cat,
                hospstay_cat,
                Persistence,
                Diabetes,
                ccschf_bl,
                Hands_only
) %>%
  tbl_summary(by = Persistence, # stratify by persistence
              statistic = list(all_continuous() ~ "{mean} ({sd})", #use mean and standard deviation for continuous variables
                               all_categorical() ~ "{n} ({p}%)"), #use count and percentage for categorical variables
              digits = list(all_categorical() ~ c(0, 2), #adds no decimal places to counts and one decimal to percentages
                            all_continuous() ~ 1), #adds one decimal place to mean and sd values
              missing = "no"
              #missing_text = "Missing (n)"
              ) %>%
  add_p(pvalue_fun = ~style_pvalue(.x, digits = 2)) %>% #adds p value
  add_overall() %>% #adds column with non-stratified summary statistics
  modify_header(label ~ "**Variable**") %>% #the asterisks bold the label
  modify_spanning_header(c("stat_1", "stat_2") ~ "**Persistence**") %>%
  bold_labels() %>%
  as_gt() %>% 
  tab_header(
    title = "Table 3. Variables associated with MRSA persistence in perexisting colonization patient"
    ) %>% 
  tab_style(cell_borders(
    sides = c("top"), style = "hidden"),
    locations = cells_title())

# View the table
mrsa_uni_c
```
## VRE Univariate_Colonization
```{r}
#| label: VRE univariate analysis_colonization
# Colonization VRE Univariate Table (Crude)
vre_pre_unique$Hands_only <- as.character(vre_pre_unique$Hands_only)

vre_pre_unique$Age <- as.numeric(vre_pre_unique$age_bl)
vre_pre_unique$cogsum_bl <- as.numeric(vre_pre_unique$cogsum_bl)
var_label(vre_pre_unique$cogsum_bl) <- "Cognitive Sum Score"

vre_uni_c <- vre_pre_unique %>%
  dplyr::select(Age,
                Sex,
                Race,
                vre_p_cat,
                vre_env_cat,
                antibiotic_bl,
                deviceuse, 
                openwound_final,
                cogsum_bl,
                hovre_bl,
                psmscore_cat,
                charlson_cat,
                hospstay_cat,
                Persistence,
                Diabetes,
                ccscopd_bl,
                Hands_only
) %>%
  tbl_summary(by = Persistence, # stratify by persistence
              statistic = list(all_continuous() ~ "{mean} ({sd})", #use mean and standard deviation for continuous variables
                               all_categorical() ~ "{n} ({p}%)"), #use count and percentage for categorical variables
              digits = list(all_categorical() ~ c(0, 2), #adds no decimal places to counts and one decimal to percentages
                            all_continuous() ~ 1), #adds one decimal place to mean and sd values
              missing = "no"
              #missing_text = "Missing (n)"
              ) %>%
  add_p(pvalue_fun = ~style_pvalue(.x, digits = 2)) %>% #adds p value
  add_overall() %>% #adds column with non-stratified summary statistics
  modify_header(label ~ "**Variable**") %>% #the asterisks bold the label
  modify_spanning_header(c("stat_1", "stat_2") ~ "**Persistence**") %>%
  bold_labels() %>%
  as_gt() %>% 
  tab_header(
    title = "Table 3. Variables associated with VRE persistence in perexisting colonization patient"
    ) %>% 
  tab_style(cell_borders(
    sides = c("top"), style = "hidden"),
    locations = cells_title())

# View the table
vre_uni_c
```
# RGNB OR NEW
```{r}
# Filter out variables with only one unique level
categorical_vars_rgnb <- c("Sex", "gnbr_p_cat", "gnbr_env_cat", "antibiotic_bl", 
                      "deviceuse", "openwound_final", "horgnb_bl", "psmscore_cat", 
                      "charlson_cat", "hospstay_cat", "Persistence", "Diabetes", 
                      "ccsdemen_bl", "Hands_only")
rgnb_acq_unique$age_bl <- as.numeric(rgnb_acq_unique$age_bl)
rgnb_acq_unique$cogsum_bl <- as.numeric(rgnb_acq_unique$cogsum_bl)

# Filter for categorical variables with at least two levels
valid_categorical_vars_rgnb <- categorical_vars_rgnb[sapply(rgnb_acq_unique[categorical_vars_rgnb], function(x) length(unique(x)) > 1)]

# Select only valid variables for analysis
rgnb_uni_or <- rgnb_acq_unique %>%
  dplyr::select(age_bl, cogsum_bl, all_of(valid_categorical_vars)) %>%
  tbl_uvregression(
    method = glm,
    y = Persistence,
    method.args = list(family = binomial), # logistic regression
    exponentiate = TRUE # to display OR instead of log(OR)
  ) %>%
  add_global_p() %>% # adds p-values for the overall test
  modify_header(
    label ~ "**Characteristic**",
    estimate ~ "**Unadjusted OR (95% CI)**",
    p.value ~ "**P value**"
  ) %>%
  bold_labels() %>%
  as_gt() %>%
  tab_header(
    title = "Table 1. Association between Participant Characteristics and Baseline MDRO Colonization or New MDRO Acquisitions"
  ) %>%
  tab_style(cell_borders(
    sides = c("top"), style = "hidden"),
    locations = cells_title()
  )

# Display the table
rgnb_uni_or
```

## RGNB Univariate_Acquisition
```{r}
#| label: RGNB univariate analysis_acquisition
# Colonization RGNB Univariate Table (Crude)
rgnb_acq_unique$Hands_only <- as.character(rgnb_acq_unique$Hands_only)

rgnb_acq_unique$Age <- as.numeric(rgnb_acq_unique$age_bl)
rgnb_acq_unique$cogsum_bl <- as.numeric(rgnb_acq_unique$cogsum_bl)
var_label(rgnb_acq_unique$cogsum_bl) <- "Cognitive Sum Score"

rgnb_uni_a <- rgnb_acq_unique %>%
  dplyr::select(Age,
                Sex,
                Race,
                gnbr_p_cat,
                gnbr_env_cat,
                antibiotic_bl,
                deviceuse, 
                openwound_final,
                cogsum_bl,
                horgnb_bl,
                psmscore_cat,
                charlson_cat,
                hospstay_cat,
                Persistence,
                Diabetes,
                ccshemi_bl,
                Hands_only
) %>%
  tbl_summary(by = Persistence, # stratify by persistence
              statistic = list(all_continuous() ~ "{mean} ({sd})", #use mean and standard deviation for continuous variables
                               all_categorical() ~ "{n} ({p}%)"), #use count and percentage for categorical variables
              digits = list(all_categorical() ~ c(0, 2), #adds no decimal places to counts and one decimal to percentages
                            all_continuous() ~ 1), #adds one decimal place to mean and sd values
              missing = "no"
              #missing_text = "Missing (n)"
              ) %>%
  add_p(pvalue_fun = ~style_pvalue(.x, digits = 2)) %>% #adds p value
  add_overall() %>% #adds column with non-stratified summary statistics
  modify_header(label ~ "**Variable**") %>% #the asterisks bold the label
  modify_spanning_header(c("stat_1", "stat_2") ~ "**Persistence**") %>%
  bold_labels() %>%
  as_gt() %>% 
  tab_header(
    title = "Table 4. Variables associated with RGNB persistence in New Acquisition patient"
    ) %>% 
  tab_style(cell_borders(
    sides = c("top"), style = "hidden"),
    locations = cells_title())

# View the table
rgnb_uni_a
```
## MRSA Univariate_Acquisition
```{r}
#| label: MRSA univariate analysis_acquisition
# Colonization MRSA Univariate Table (Crude)
mrsa_acq_unique$Hands_only <- as.character(mrsa_acq_unique$Hands_only)

mrsa_acq_unique$Age <- as.numeric(mrsa_acq_unique$age_bl)
mrsa_acq_unique$cogsum <- as.numeric(mrsa_acq_unique$cogsum_bl)
var_label(mrsa_acq_unique$cogsum) <- "Cognitive Sum Score"

mrsa_uni_a <- mrsa_acq_unique %>%
  dplyr::select(Age,
                Sex,
                Race,
                mrsa_p_cat,
                mrsa_env_cat,
                antibiotic_bl,
                deviceuse, 
                openwound_final,
                cogsum,
                homrsa_bl,
                psmscore_cat,
                charlson_cat,
                hospstay_cat,
                Persistence,
                Diabetes,
                ccschf_bl,
                Hands_only
) %>%
  tbl_summary(by = Persistence, # stratify by persistence
              type = list(cogsum ~ "continuous"), # explicitly set cogsum as continuous
              statistic = list(all_continuous() ~ "{mean} ({sd})", #use mean and standard deviation for continuous variables
                               all_categorical() ~ "{n} ({p}%)"), #use count and percentage for categorical variables
              digits = list(all_categorical() ~ c(0, 2), #adds no decimal places to counts and one decimal to percentages
                            all_continuous() ~ 1), #adds one decimal place to mean and sd values
              missing = "no"
              #missing_text = "Missing (n)"
              ) %>%
  add_p(pvalue_fun = ~style_pvalue(.x, digits = 2)) %>% #adds p value
  add_overall() %>% #adds column with non-stratified summary statistics
  modify_header(label ~ "**Variable**") %>% #the asterisks bold the label
  modify_spanning_header(c("stat_1", "stat_2") ~ "**Persistence**") %>%
  bold_labels() %>%
  as_gt() %>% 
  tab_header(
    title = "Table 4. Variables associated with MRSA persistence in New Acquisition patient"
    ) %>% 
  tab_style(cell_borders(
    sides = c("top"), style = "hidden"),
    locations = cells_title())

# View the table
mrsa_uni_a
```
## VRE Univariate_Acquisition
```{r}
#| label: VRE univariate analysis_acquisition
vre_acq_unique$Hands_only <- as.character(vre_acq_unique$Hands_only)

vre_acq_unique$Age <- as.numeric(vre_acq_unique$age_bl)
vre_acq_unique$cogsum <- as.numeric(vre_acq_unique$cogsum_bl)
var_label(vre_acq_unique$cogsum) <- "Cognitive Sum Score"

vre_uni_a <- vre_acq_unique %>%
  dplyr::select(Age,
                Sex,
                Race,
                vre_p_cat,
                vre_env_cat,
                antibiotic_bl,
                deviceuse, 
                openwound_final,
                cogsum,
                hovre_bl,
                psmscore_cat,
                charlson_cat,
                hospstay_cat,
                Persistence,
                Diabetes,
                ccschf_bl,
                Hands_only
) %>%
  tbl_summary(by = Persistence, # stratify by persistence
              type = list(cogsum ~ "continuous"), # explicitly set cogsum as continuous
              statistic = list(all_continuous() ~ "{mean} ({sd})", #use mean and standard deviation for continuous variables
                               all_categorical() ~ "{n} ({p}%)"), #use count and percentage for categorical variables
              digits = list(all_categorical() ~ c(0, 2), #adds no decimal places to counts and one decimal to percentages
                            all_continuous() ~ 1), #adds one decimal place to mean and sd values
              missing = "no"
              #missing_text = "Missing (n)"
              ) %>%
  add_p(pvalue_fun = ~style_pvalue(.x, digits = 2)) %>% #adds p value
  add_overall() %>% #adds column with non-stratified summary statistics
  modify_header(label ~ "**Variable**") %>% #the asterisks bold the label
  modify_spanning_header(c("stat_1", "stat_2") ~ "**Persistence**") %>%
  bold_labels() %>%
  as_gt() %>% 
  tab_header(
    title = "Table 4. Variables associated with VRE persistence in New Acquisition patient"
    ) %>% 
  tab_style(cell_borders(
    sides = c("top"), style = "hidden"),
    locations = cells_title())

# View the table
vre_uni_a
```

# Table 5: Multivariate for Persistence
## Stepwise - Colonization
```{r}
#| label: RGNB_Colonization
# Multivariate logistic models
# Load necessary libraries
library(MASS)
library(dplyr)

# Define the variables to check for missing values
variables_to_check <- c("age_bl", "BMI", "gender", "race", "cogsum_bl", "deviceuse", "charlson_cat", 
                        "horgnb_bl", "antibiotic_bl", "openwound_final", "hospstay")

# Subset the data to remove rows with missing values in the specified variables
rgnb_unique_c_complete <- rgnb_pre_unique[complete.cases(rgnb_unique_c[, variables_to_check]), ]

# Define the initial full model with all potential risk factors for RGNB
rgnb_full_model_c <- glm(persistence ~ age_bl + BMI + gender + race + cogsum_bl + deviceuse + charlson_cat + horgnb_bl + antibiotic_bl + openwound_final + hospstay + psmscore,
                        data = rgnb_unique_c_complete, 
                        family = binomial)

# Perform stepwise regression for MRSA
rgnb_stepwise_model_c <- stepAIC(rgnb_full_model_c, direction = "both", trace = FALSE)

# Summary of the final MRSA model
summary(rgnb_stepwise_model_c)

# Extract and print final coefficients for MRSA
rgnb_final_coefficients <- summary(rgnb_stepwise_model_c)$coefficients
print(rgnb_final_coefficients)
```
## Multivariate - New Acquisition
```{r}
#| label: multivariate analysis_New acquisition
# Multivariate logistic models
model_rgnb_a <- glm(persistence ~ age_bl + gender + deviceuse + urincath + ccshemi_bl + ccsmsrd_bl, 
                   data = rgnb_unique_a, 
                   family = binomial(link = "logit"))
summary(model_rgnb_a)

model_mrsa_a <- glm(persistence ~ age_bl + gender + hospstay, 
                   data = mrsa_unique_a, 
                   family = binomial(link = "logit"))
summary(model_mrsa_a)

model_vre_a <- glm(persistence ~ age_bl + BMI + gender + vre_p + ccschf_bl, 
                   data = vre_unique_a, 
                   family = binomial(link = "logit"))
summary(model_vre_a)
```

# Figure 3.Survival analysis
## New Acquisition
```{r}
# Combine persistence, duration, and clearance into a unified format
combined_long <- combined_unique_a %>%
  mutate(
    Persistence = case_when(
      organism == "RGNB" ~ rgnb_acq_persistence,
      organism == "MRSA" ~ mrsa_acq_persistence,
      organism == "VRE" ~ vre_acq_persistence
    ),
    Duration = case_when(
      organism == "RGNB" ~ rgnb_acq_duration,
      organism == "MRSA" ~ mrsa_acq_duration,
      organism == "VRE" ~ vre_acq_duration
    ),
    Clearance = case_when(
      organism == "RGNB" ~ rgnb_acq_cleared,
      organism == "MRSA" ~ mrsa_acq_cleared,
      organism == "VRE" ~ vre_acq_cleared
    )
  )

combined_long <- combined_long %>%
  filter(mrsa_acq_totvisits_r > 5 |
         rgnb_acq_totvisits_r > 5 |
         vre_acq_totvisits_r > 5)

# Filter for persistence data
persistence_a <- combined_long %>%
  filter(Persistence == "Persistent")

persistence_a$Duration <- as.numeric(persistence_a$Duration)

# Create a survival object
surv_a <- Surv(time = persistence_a$Duration, 
               event = persistence_a$Clearance == "Yes")

# Fit the Kaplan-Meier model
km_curve_a <- survfit(surv_a ~ organism, data = persistence_a)
summary(km_curve_a)
survdiff(surv_a ~ organism, data = persistence_a)

# Plot the Kaplan-Meier curve
gg_surv_a <- ggsurvplot(km_curve_a, 
           data = persistence_a,
           pval = TRUE,
           pval.method = TRUE,
           pval.coord = c(0, 0.1), # position the p-value on the plot
           pval.method.coord = c(0, 0.15), # position the method p-value on the plot
           ggtheme = theme_minimal(),
           # risk.table = TRUE,       # To show the risk table
           # risk.table.col = "strata",# Change risk table color by groups
           legend.labs = c("MRSA", "RGNB", "VRE"), # Change group labels
           legend.title = "Organism",
           xlab = "Clearance of colonization (days)",
           ylab = "Probability of MDROs Persistence",
           title = "Kaplan-Meier Survival Curve for New Acquisition Residents"
)

# Print the plot
print(gg_surv_a)

# Save the plot by extracting the ggplot component
ggsave(file = "survival_1.png", plot = gg_surv_a$plot, width = 8, height = 6, dpi = 600)


# Calculate median survival times
median_surv_a <- surv_median(km_curve_a)
median_surv_a

# Add median survival times to the plot
# medians <- median_surv$median
#for(i in 1:nrow(median_surv)) {
#  gg_surv$plot <- gg_surv$plot +
#    annotate("text", x = medians[i], y = 0.5, label = #paste("Median:", round(medians[i], 2)),
#             color = "black", vjust = -1) # Adjust color and vjust as needed
#}

# Extract survival probabilities at specific times (e.g., 50 and 100 days)
# surv_probs <- summary(km_curve_a, times = c(50, 100))
# print(surv_probs)

# Add survival probabilities to the plot
# for (i in 1:length(km_curve_a$strata)) {
#  organism_name <- names(km_curve_a$strata)[i]
#  prob_50 <- surv_probs$surv[which(surv_probs$strata == organism_name & surv_probs$time == 50)]
#  prob_100 <- surv_probs$surv[which(surv_probs$strata == organism_name & surv_probs$time == 100)]

#  gg_surv$plot <- gg_surv$plot +
 #   annotate("text", x = 50, y = prob_50, label = paste(organism_name, "50 days:", round(prob_50, 2)),
  #           color = "black", vjust = -1, hjust = 0) +
   # annotate("text", x = 100, y = prob_100, label = paste(organism_name, "100 days:", round(prob_100, 2)),
    #         color = "black", vjust = -1, hjust = 0)

# Load cowplot library
library(cowplot)

# Assemble the components manually
combined_plot <- cowplot::plot_grid(
     gg_surv_a$plot,
     gg_surv_a$table,
     ncol = 1,
     rel_heights = c(3, 1)  # Adjust the heights as needed
 )
 
 # Save the combined plot
ggsave(filename = "survival_1.png", plot = combined_plot, width = 8, height = 6, dpi = 600)
#}
```
### Split Acquisition
```{r}
# Create subsets for each organism
persistence_rgnb_a <- subset(persistence_a, organism == "RGNB")
persistence_mrsa_a <- subset(persistence_a, organism == "MRSA")
persistence_vre_a <- subset(persistence_a, organism == "VRE")

# Kaplan-Meier analysis for RGNB
surv_rgnb_a <- Surv(time = persistence_rgnb_a$duration, event = persistence_rgnb_a$cleared == "Yes")
km_curve_rgnb_a <- survfit(surv_rgnb_a ~ 1)
summary(km_curve_rgnb_a)
ggsurvplot(km_curve_rgnb_a, 
           data = persistence_rgnb_a,
           ggtheme = theme_minimal(),
           xlab = "Clearance of colonization (days)",
           ylab = "Probability of MDROs Persistence",
           title = "Kaplan-Meier Survival Curve for RGNB Residents",
           legend = "none"
)

# Kaplan-Meier analysis for MRSA
surv_mrsa_a <- Surv(time = persistence_mrsa_a$duration, event = persistence_mrsa_a$cleared == "Yes")
km_curve_mrsa_a <- survfit(surv_mrsa_a ~ 1)
summary(km_curve_mrsa_a)
ggsurvplot(km_curve_mrsa_a, 
           data = persistence_mrsa_a,
           ggtheme = theme_minimal(),
           xlab = "Clearance of colonization (days)",
           ylab = "Probability of MDROs Persistence",
           title = "Kaplan-Meier Survival Curve for MRSA Residents",
           legend = "none"
)

# Kaplan-Meier analysis for VRE
surv_vre_a <- Surv(time = persistence_vre_a$duration, event = persistence_vre_a$cleared == "Yes")
km_curve_vre_a <- survfit(surv_vre_a ~ 1)
summary(km_curve_vre_a)
ggsurvplot(km_curve_vre_a, 
           data = persistence_vre_a,
           ggtheme = theme_minimal(),
           xlab = "Clearance of colonization (days)",
           ylab = "Probability of MDROs Persistence",
           title = "Kaplan-Meier Survival Curve for VRE Residents",
           legend = "none"
)

# Cox proportional hazards model for RGNB
cox_rgnb_a <- coxph(surv_rgnb_a ~ age_bl + gender, data = persistence_rgnb_a)
summary(cox_rgnb_a)
exp(cbind(coef(cox_rgnb_a), confint(cox_rgnb_a)))
cox.zph(cox_rgnb_a)
plot(cox.zph(cox_rgnb_a))

# Cox proportional hazards model for VRE
cox_vre_a <- coxph(surv_vre_a ~ age_bl + gender, data = persistence_vre_a)
summary(cox_vre_a)
exp(cbind(coef(cox_vre_a), confint(cox_vre_a)))
cox.zph(cox_vre_a)
plot(cox.zph(cox_vre_a))
```

## Colonization
```{r}
# Load necessary libraries
library(survival)
library(ranger)
library(ggplot2)
library(dplyr)
library(ggfortify)
library(survminer)

# Combine persistence, duration, and clearance into a unified format
combined_long_c <- combined_unique_c %>%
  mutate(
    Persistence = case_when(
      organism == "RGNB" ~ rgnb_pre_persistence,
      organism == "MRSA" ~ mrsa_pre_persistence,
      organism == "VRE" ~ vre_pre_persistence
    ),
    Duration = case_when(
      organism == "RGNB" ~ rgnb_pre_duration,
      organism == "MRSA" ~ mrsa_pre_duration,
      organism == "VRE" ~ vre_pre_duration
    ),
    Clearance = case_when(
      organism == "RGNB" ~ rgnb_pre_cleared,
      organism == "MRSA" ~ mrsa_pre_cleared,
      organism == "VRE" ~ vre_pre_cleared
    )
  )

# Filter for persistence data
combined_long_c <- combined_long_c %>%
  filter(mrsa_pre_totvisits_r > 5 |
         rgnb_pre_totvisits_r > 5 |
         vre_pre_totvisits_r > 5)

persistence_c <- combined_long_c %>%
  filter(Persistence == "Persistent")

persistence_c$Duration <- as.numeric(persistence_c$Duration)

# Create a survival object
surv_c <- Surv(time = persistence_c$Duration, 
             event = persistence_c$Clearance == "Yes")

# Fit the Kaplan-Meier model
km_curve_c <- survfit(surv_c ~ organism, data = persistence_c)
summary(km_curve_c)
survdiff(surv_c ~ organism, data = persistence_c)

# Plot the Kaplan-Meier curve
gg_surv_c <- ggsurvplot(km_curve_c, 
           data = persistence_c,
           pval = TRUE,
           pval.method = TRUE,
           pval.coord = c(0, 0.1), # position the p-value on the plot
           pval.method.coord = c(0, 0.15), # position the method p-value on the plot
           ggtheme = theme_minimal(),
           # conf.int = TRUE,
           # risk.table = TRUE,       # To show the risk table
           # risk.table.col = "strata",# Change risk table color by groups
           legend.labs = c("MRSA", "RGNB", "VRE"), # Change group labels
           legend.title = "Organism",
           xlab = "Clerance of colonization (days)",
           ylab = "Probability of MDROs Persistnece",
           title = "Kaplan-Meier Survival Curve for Preexisting Colonization Residents"
)

# Calculate median survival times
median_surv_c <- surv_median(km_curve_c)
median_surv_c

# Add median survival times to the plot
# medians <- median_surv$median
#for(i in 1:nrow(median_surv)) {
#  gg_surv$plot <- gg_surv$plot +
#    annotate("text", x = medians[i], y = 0.5, label = #paste("Median:", round(medians[i], 2)),
#             color = "black", vjust = -1) # Adjust color and vjust as needed
#}

# Extract survival probabilities at specific times (e.g., 50 and 100 days)
# surv_probs <- summary(km_curve_a, times = c(50, 100))
# print(surv_probs)

# Add survival probabilities to the plot
# for (i in 1:length(km_curve_a$strata)) {
#  organism_name <- names(km_curve_a$strata)[i]
#  prob_50 <- surv_probs$surv[which(surv_probs$strata == organism_name & surv_probs$time == 50)]
#  prob_100 <- surv_probs$surv[which(surv_probs$strata == organism_name & surv_probs$time == 100)]

#  gg_surv$plot <- gg_surv$plot +
 #   annotate("text", x = 50, y = prob_50, label = paste(organism_name, "50 days:", round(prob_50, 2)),
  #           color = "black", vjust = -1, hjust = 0) +
   # annotate("text", x = 100, y = prob_100, label = paste(organism_name, "100 days:", round(prob_100, 2)),
    #         color = "black", vjust = -1, hjust = 0)
#}

# Print the plot
print(gg_surv_c)
# Load cowplot library
library(cowplot)
 
# Assemble the components manually
combined_plot_c <- cowplot::plot_grid(gg_surv_c$plot, 
                                    gg_surv_c$table,
                                    ncol = 1,
                                    rel_heights = c(3, 1)  # Adjust the heights as needed
                                    )

# Save the combined plot
ggsave(filename = "survival_2.png", plot = combined_plot_c, width = 8, height = 6, dpi = 600)

# Cox proportional hazard
cox_c <- coxph(surv_c ~ organism, data = persistence_c)
summary(cox_c)
# Hazard Ratios (95% CI)
exp(cbind(coef(cox_c),coefci(cox_c)))
cox.zph(cox_c) # tests whether the hazards are proportional each variable separately
plot(cox.zph(cox_c)) # Check assumptions
```
### Split Colonization
```{r}
persistence_rgnb_c <- subset(persistence_c, organism == "RGNB")
persistence_mrsa_c <- subset(persistence_c, organism == "MRSA")
persistence_vre_c <- subset(persistence_c, organism == "VRE")

# Kaplan-Meier analysis for RGNB
surv_rgnb_c <- Surv(time = persistence_rgnb_c$duration, event = persistence_rgnb_c$cleared == "Yes")
km_curve_rgnb_c <- survfit(surv_rgnb_c ~ 1)
summary(km_curve_rgnb_c)
ggsurvplot(km_curve_rgnb_c, 
           data = persistence_rgnb_c,
           ggtheme = theme_minimal(),
           xlab = "Clearance of colonization (days)",
           ylab = "Probability of MDROs Persistence",
           title = "Kaplan-Meier Survival Curve for RGNB Residents",
           legend = "none"
)

# Kaplan-Meier analysis for MRSA
surv_mrsa_c <- Surv(time = persistence_mrsa_c$duration, event = persistence_mrsa_c$cleared == "Yes")
km_curve_mrsa_c <- survfit(surv_mrsa_c ~ 1)
summary(km_curve_mrsa_c)
ggsurvplot(km_curve_mrsa_c, 
           data = persistence_mrsa_c,
           ggtheme = theme_minimal(),
           xlab = "Clearance of colonization (days)",
           ylab = "Probability of MDROs Persistence",
           title = "Kaplan-Meier Survival Curve for MRSA Residents",
           legend = "none"
)

# Kaplan-Meier analysis for VRE
surv_vre_c <- Surv(time = persistence_vre_c$duration, event = persistence_vre_c$cleared == "Yes")
km_curve_vre_c <- survfit(surv_vre_c ~ 1)
summary(km_curve_vre_c)
ggsurvplot(km_curve_vre_c, 
           data = persistence_vre_c,
           ggtheme = theme_minimal(),
           xlab = "Clearance of colonization (days)",
           ylab = "Probability of MDROs Persistence",
           title = "Kaplan-Meier Survival Curve for VRE Residents",
           legend = "none"
)

# Cox proportional hazards model for RGNB
cox_rgnb_c <- coxph(surv_rgnb_c ~ age_bl + gender, data = persistence_rgnb_c)
summary(cox_rgnb_c)
exp(cbind(coef(cox_rgnb_c), confint(cox_rgnb_c)))
cox.zph(cox_rgnb_c)
plot(cox.zph(cox_rgnb_c))

# Cox proportional hazards model for MRSA
cox_mrsa_c <- coxph(surv_mrsa_c ~ age_bl + gender, data = persistence_mrsa_c)
summary(cox_mrsa_c)
exp(cbind(coef(cox_mrsa_c), confint(cox_mrsa_c)))
cox.zph(cox_mrsa_c)
plot(cox.zph(cox_mrsa_c))

# Cox proportional hazards model for VRE
cox_vre_c <- coxph(surv_vre_c ~ age_bl + gender, data = persistence_vre_c)
summary(cox_vre_c)
exp(cbind(coef(cox_vre_c), confint(cox_vre_c)))
cox.zph(cox_vre_c)
plot(cox.zph(cox_vre_c))
```



